---
title: "Comparison of egg-adult viability between Ace genotypes in the absence of malathion"
author: "Marianna Karageorgi"
date: "2025-05-07"
---

# 1. Data

Starting July 6th 2023, eggs of each D.melanogaster genotype (SS, R1R1, R2R2, R3R3, SR1, SR2, SR3) were exposed to regular Drosophila media (50 eggs/trial). For each genotype we had 10 replicates.

Ace-S: IGFG, Ace-R1: [V]{.underline}GFG, Ace-R2: [V]{.underline}GF[A]{.underline}, Ace-R3: [VAY]{.underline}G

### Data Wrangling

```{r}
library(tidyverse)
library(data.table)
```

```{r}
data = fread("./data/Viability_FINAL.csv")

data.viability = data %>%
  mutate(
    Emerge = rowSums(select(., 4:12), na.rm = TRUE),
    Total = 50,
    Not_Emerge = Total - Emerge,
    Survival = Emerge / Total * 100) %>%
  
    mutate(Genotype = factor(Genotype, 
                             levels = c("SS","SR1","SR2","SR3","R1R1", "R2R2", "R3R3"))) %>%
  select(Genotype, Emerge, Not_Emerge, Survival, Total)
```

### Summary Statistics

```{r}
data.viability.summary = group_by(data.viability, Genotype) %>%
  summarise(
    count = n(),
    mean = mean(Survival, na.rm = TRUE),
    sd = sd(Survival, na.rm = TRUE),
    se = sd / sqrt(10)
  )
```

### Viability data: Exploratory Data Analysis

```{r}
library("ggpubr")
library("ggplot2")


# Define the custom colors for the palette
genotype_colors = c("black", "#FCCC1A","#8601AF","#FE2712",  "#FCCC1A", "#8601AF",   "#FE2712")

# Define the linetypes for the boxes
linetypes = c(1, 2, 2, 2, 1, 1, 1)


viability_plot = ggplot(data.viability, 
                        aes(Genotype, Survival)) +
  
  geom_boxplot(aes(linetype = Genotype, color = Genotype), fill = NA) +
  
  geom_point(aes(color = Genotype)) +  
  scale_color_manual(values = genotype_colors) +
  scale_linetype_manual(values = linetypes) +
  labs(x = NULL, y = "Egg to adult vialibity (%)") +
  ylim(40, 100) + 
  theme_classic() 

viability_plot 
```

# 2. Model data

Use logistic regression model to analyze viability data for homozygous sensitive SS, homozygous resistant RR, and heterozygous SR genotypes. Use a logistic model because of grouped binary data (nr of egg emerged out of 50 eggs).

The model is:

logit(p)=log(p/(1-p)= b0​+b1​⋅SR+b2​⋅RR+ϵ

The odds of emergence for each genotype is represented:

For SS (reference group): odds.emerge.SS​=b0​

For SR: odds.emerge.SR=b0+b1

For RR: odds.emerge.RR=b0+b2

Transform logit values to probabilities of emergence:

proportion_SS = exp(logit_SS) / (1 + exp(logit_SS))

proportion_SR = exp(logit_SR) / (1 + exp(logit_SR))

proportion_RR = exp(logit_RR) / (1 + exp(logit_RR))

```{r}
# Fit logistic binomial model
model.R = glm(cbind(Emerge, Not_Emerge) ~ Genotype, data = data.viability, family = "binomial")

summary(model.R)
```

# 3. Metrics & Bootstrap for viability

Estimate difference and metric parameters using coefficients from the logistic regression model. Then use bootstrapping to estimate confidence intervals for each parameter.

**Additivity: Comparison between SR vs. (SS+RR)/2**

Additivity is calculated as the difference between the SR genotype and the average of SS and RR genotypes. It can be compared to the expected value under additivity.

-   Expected value under additivity:

    SR- (SS+RR)/2 = proportion_SR -(proportion_SS+proportion_RR)/2

-   Significance testing with two one-sided t-test: H0 = the estimate for additivity is 0. Ha1 = the estimate for dominance is higher than 0 Ha2 = the estimate for dominance is lower than 0.

**Selection coefficient: Estimate ratio (RR - SS)/ SS**

The selection coefficient is the ratio of the difference between RR and SS genotypes to the SS genotype. This parameter indicates how selection affects the fitness of the resistant genotype compared to the sensitive genotype.

-   Estimate ratio (RR-SS)/SS = (proportion_RR-proportion_SS)/proportion_SS

-   Significance testing with two-sided t-test: H0 = the estimate for selection equals to 0. Ha = the estimate for selection is different from 0.

**Dominance coefficient: Estimate (SR−SS)/(RR−SS) (h)**

The dominance coefficient measures the degree to which the heterozygous genotype (SR) exhibits fitness that is intermediate between the homozygous genotypes (SS and RR).

-   Estimate ratio (proportion_SR-proportion_SS)/(proportion_RR-proportion_SS)

-   Significance testing with two one-sided t-test: H0 = the estimate for dominance is 0.5. Ha1 = the estimate for dominance is higher than 0.5. Ha2 = the estimate for dominance is lower than 0.5

```{r}
library(boot)
library(broom)

# bootstrap & estimate metrics

# bootstrap function
metrics.boot = function(data, original) { 
  # Create bootstrap sample
  d = data %>%
    group_by(Genotype) %>%
    # randomly select 10 rows from each Genotype group, sampling with replacement
    slice_sample(n = 10, replace = TRUE) %>% # n = number of reps in each fecundity group
    ungroup()
  
  # Fit model to bootstrap sample 
  model = glm(cbind(Emerge, Not_Emerge) ~ Genotype, data = d, family = "binomial")
  coefficients = coef(model)
  
  # Extract coefficients to calculate odds of emergence for each genotype 
  coefficients = model$coefficients  # Fix: extract coefficients from model
  b0 = coefficients["(Intercept)"]
  b1 = coefficients["GenotypeSR1"]
  b2 = coefficients["GenotypeR1R1"]
  b3 = coefficients["GenotypeSR2"]
  b4 = coefficients["GenotypeR2R2"]
  b5 = coefficients["GenotypeSR3"]
  b6 = coefficients["GenotypeR3R3"]
  
  # Transform logit values to probabilities of emergence for each genotype to calculate metrics for each boot
  ss    = exp(b0) / (1 + exp(b0))
  sr1   = exp(b0 + b1) / (1 + exp(b0 + b1))
  r1r1  = exp(b0 + b2) / (1 + exp(b0 + b2))
  sr2   = exp(b0 + b3) / (1 + exp(b0 + b3))
  r2r2  = exp(b0 + b4) / (1 + exp(b0 + b4))
  sr3   = exp(b0 + b5) / (1 + exp(b0 + b5))
  r3r3  = exp(b0 + b6) / (1 + exp(b0 + b6))
  
  
  # Calculate all metrics for each R allele for each boot
  # For R1
  ss_sr1_diff = ss - sr1  # Difference between SS and SR1
  ss_r1r1_diff = ss - r1r1  # Difference between SS and R1R1
  sr1_r1r1_diff = sr1 - r1r1  # Difference between SR1 and R1R1 (CORRECTED)
  dev.add.r1 = sr1 - (ss + r1r1)/2  # Deviation from additivity
  s.r1 = (r1r1 - ss)/ss  # Selection coefficient
  h.r1 = (sr1 - ss)/(r1r1 - ss)  # Dominance coefficient 
  
  
  # For R2
  ss_sr2_diff = ss - sr2
  ss_r2r2_diff = ss - r2r2
  sr2_r2r2_diff = sr2 - r2r2
  dev.add.r2 = sr2 - (ss + r2r2)/2
  s.r2 = (r2r2 - ss)/ss
  h.r2 = (sr2 - ss)/(r2r2 - ss)
 
  
  # For R3
  ss_sr3_diff = ss - sr3
  ss_r3r3_diff = ss - r3r3
  sr3_r3r3_diff = sr3 - r3r3
  dev.add.r3 = sr3 - (ss + r3r3)/2
  s.r3 = (r3r3 - ss)/ss
  h.r3 = (sr3 - ss)/(r3r3 - ss) 
  
  # Return with explicit names
  result_vector = c(r1r1, r2r2, r3r3, sr1, sr2, sr3, ss, 
                   dev.add.r1, dev.add.r2, dev.add.r3,
                   s.r1, s.r2, s.r3,
                   h.r1, h.r2, h.r3,
                   ss_sr1_diff, ss_sr2_diff, ss_sr3_diff,
                   ss_r1r1_diff, ss_r2r2_diff, ss_r3r3_diff,
                   sr1_r1r1_diff, sr2_r2r2_diff, sr3_r3r3_diff)
  
  names(result_vector) = c("R1R1", "R2R2", "R3R3","SR1", "SR2", "SR3", "SS",
                          "DevAdd_R1", "DevAdd_R2", "DevAdd_R3",
                          "S_R1", "S_R2", "S_R3",
                          "H_R1", "H_R2", "H_R3",
                          "diff_SS.SR1", "diff_SS.SR2", "diff_SS.SR3",
                          "diff_SS.R1R1", "diff_SS.R2R2", "diff_SS.R3R3",
                          "diff_SR1.R1R1", "diff_SR2.R2R2", "diff_SR3.R3R3")
  return(result_vector)
}

# run the bootstrap
set.seed(123)
nboot=1000
result = boot(data = data.viability,
               statistic = metrics.boot,
               R = nboot)  # R for number of bootstraps

result$t0 # original estimate based on data
result$t  # boot samples estimates

summary(result)
tidy(result)
```

```{r}
### Viability ###

# Report the original estimate (result$t0) based on the data  
# Use the 2.5% and 97.5% quantiles of the bootstrap distribution (result$t) for the 95% ci

# Process bootstrap results into a single LD50 table
viability.results = function(result) {
  # Define genotypes and indices
  genotypes = c("R1R1", "R2R2", "R3R3", "SR1", "SR2", "SR3", "SS")
  viability_indices = 1:7
  
  # Process fecundity values
  viability_table = do.call(rbind, lapply(1:length(genotypes), function(i) {
    # Extract values
    original = result$t0[i]
    boot_samples = result$t[, i]
    ci = quantile(boot_samples, c(0.025, 0.975))
    
    data.frame(
      Genotype = genotypes[i],
      Estimate = original,
      Lower.CI = ci[1],
      Upper.CI = ci[2]
    )
  }))
  
  return(viability_table)
}

# Get the viability data frame
viability = viability.results(result)
viability
```

```{r}
### DIFFERENCES & METRICS ####

# Report the original estimate (result$t0) based on the data  
# Use the 2.5% and 97.5% quantiles of the bootstrap distribution (result$t) for the 95% ci
# Report the p-value calculated from the bootstrap distribution

# First, define the p.value function
p.value = function(boot_samples, null_value) {
  p_less = mean(boot_samples <= null_value, na.rm = TRUE)
  p_greater = mean(boot_samples >= null_value, na.rm = TRUE)
  return(2 * min(p_less, p_greater)) # which tail is more extreme
}

# Then define metrics.results
metrics.results = function(result) {
  # Define alleles
  alleles = c("R1", "R2", "R3")
  
  # Define all metrics
  all_metrics = list(
    
    # Derived metrics
    list(type = "DevAdd", name = "Additivity", indices = 8:10, null_value = 0),
    list(type = "S", name = "Selection coefficient", indices = 11:13, null_value = 0),
    list(type = "H", name = "Dominance coefficient", indices = 14:16, null_value = 0.5),
    
    # Derived differences
    list(type = "diff_SS.SR", name = "SS-SR difference", indices = 17:19, null_value = 0),
    list(type = "diff_SS.RR", name = "SS-RR difference", indices = 20:22, null_value = 0),
    list(type = "diff_SR.RR", name = "SR-RR difference", indices = 23:25, null_value = 0)
  )
  
  # Process metrics
  metrics_table = do.call(rbind, lapply(1:length(all_metrics), function(m) {
    metric = all_metrics[[m]]
    
    do.call(rbind, lapply(1:length(alleles), function(a) {
      idx = metric$indices[a]
      original = result$t0[idx]
      boot_samples = result$t[, idx]
      ci = quantile(boot_samples, c(0.025, 0.975), na.rm = TRUE)
      p_val = p.value(boot_samples, metric$null_value)
      
      data.frame(
        Metric = metric$name,
        Allele = alleles[a],
        Estimate = original,
        Lower.CI = ci[1],
        Upper.CI = ci[2],
        p.value = p_val
      )
    }))
  }))
  
  # Return the formatted table
  return(metrics_table)
}

# Now call metrics.results
metrics = metrics.results(result)
metrics[metrics$p.value==0,]$p.value= 1/nboot # this is so that p.value is not 0
```

# 3. Table for viability

```{r}
# combine dataframes for the final table
viability.format = viability %>%
  rename(Genotype.Allele = Genotype) %>%
  mutate(Metric = "viability") %>%  
  dplyr::select(Genotype.Allele, Metric, Estimate, Lower.CI, Upper.CI) %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  select(Metric, Genotype.Allele, Estimate, Lower.CI, Upper.CI)

metrics.format =  metrics %>%
  rownames_to_column(var = "column") %>% separate(column, into = c("Metric", "Genotype.Allele"), sep = "_") %>%
  dplyr::select(-Allele) %>%
  mutate(across(where(is.numeric), ~round(., 3)))

all.metrics = bind_rows(viability.format, metrics.format) %>%
  mutate(Genotype.Allele = factor(Genotype.Allele, 
                                  levels = c("SS", "SR1", "SR2", "SR3", "R1R1", "R2R2", "R3R3", 
                                             "SS.SR1","SS.SR2","SS.SR3", "SS.R1R1","SS.R2R2","SS.R3R3", "SR1.R1R1","SR2.R2R2","SR3.R3R3",
                                             "R1", "R2", "R3")),
         Metric = factor(Metric, 
                         levels = c("viability","diff","DevAdd", "S", "H"))) %>%
  arrange(Metric, Genotype.Allele)
```

```{r}
library(kableExtra)
library(htmltools)
library(webshot)
```

```{r}
all.metrics.table = all.metrics %>%
  unite(col = "Parameter", Genotype.Allele, Metric, sep = "_") %>%
  rename(Viability = Estimate) %>%
  
  mutate(Parameter = case_when(
    
    Parameter == "SS_viability" ~ "SS",
    Parameter == "SR1_viability" ~ "SR1",
    Parameter == "R1R1_viability" ~ "R1R1",
    Parameter == "SR2_viability" ~ "SR2",
    Parameter == "R2R2_viability" ~ "R2R2",
    Parameter == "SR3_viability" ~ "SR3",
    Parameter == "R3R3_viability" ~ "R3R3",

    Parameter == "SS.SR1_diff" ~ "SS vs. SR1",
    Parameter == "SS.R1R1_diff" ~ "SS vs. R1R1",
    Parameter == "SR1.R1R1_diff" ~ "SR1 vs. R1R1",
    Parameter == "SS.SR2_diff" ~ "SS vs. SR2",
    Parameter == "SS.R2R2_diff" ~ "SS vs. R2R2",
    Parameter == "SR2.R2R2_diff" ~ "SR2 vs. R2R2",
    Parameter == "SS.SR3_diff" ~ "SS vs. SR3",
    Parameter == "SS.R3R3_diff" ~ "SS vs. R3R3",
    Parameter == "SR3.R3R3_diff" ~ "SR3 vs. R3R3",

    Parameter == "R1_DevAdd" ~ "SR1 - (SS + R1R1)/2 (additivity)",
    Parameter == "R2_DevAdd" ~ "SR2 - (SS + R2R2)/2 (additivity)",
    Parameter == "R3_DevAdd" ~ "SR3 - (SS + R3R3)/2 (additivity)",

    Parameter == "R1_S" ~ "(R1R1 - SS) / SS (selection, s)",
    Parameter == "R2_S" ~ "(R2R2 - SS) / SS (selection, s)",
    Parameter == "R3_S" ~ "(R3R3 - SS) / SS (selection, s)",

    Parameter == "R1_H" ~ "(SR1 - SS) / (R1R1 - SS) (dominance, h)",
    Parameter == "R2_H" ~ "(SR2 - SS) / (R2R2 - SS) (dominance, h)",
    Parameter == "R3_H" ~ "(SR3 - SS) / (R3R3 - SS) (dominance, h)")) %>%
   #replace NA with -
  mutate(across(everything(), ~replace(., is.na(.), "-")))


table.kable = kable(x = all.metrics.table, digits = 3, align = 'l',
                       caption = "<b>Effect estimates from logistic model analysis for viability</b>",
                    row.names = FALSE) %>%
                       kable_classic(full_width = TRUE, html_font = "Times New Roman", font_size = 16) %>%
                       kable_styling(latex_options = c("striped", "hold_position"), position = "center") %>%
                       row_spec(0, bold = TRUE, font_size = 16) %>%  # Make header bold and set font size
                       row_spec(1:nrow(all.metrics.table), font_size = 16)  # Smaller font for values



table.kable

# Save the table as an HTML file
save_kable(table.kable, file = "./tables/table_viability.html")

# Use webshot to convert HTML to PDF
webshot("./tables/table_viability.html", file = "./tables/table_viability.pdf")
```

## Data for simulations

```{r}
table.viability = all.metrics.table%>%
  
  filter(Parameter %in% c("SS", "SR1","SR2", "SR3", "R1R1", "R2R2", "R3R3")) %>%
  
  dplyr::select(Parameter, Viability)

write_csv(table.viability, "./data/model.viability.0ppm.csv")
```

# 4. Plots

### 4.1 Ace R1 genotype

```{r}
genotype_labels = c("SS","SR1","R1R1")

# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#30a4bc", "#ffce1b")
genotype_colors = c("#165094", "#7396BF","#ce8e58")

# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)



R1_viability = ggplot() +
  
  annotate("segment", x = 1, xend = 3, 
           y = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "viability") %>% pull(Estimate)*100, 
            yend = all.metrics %>% filter(Genotype.Allele == "R1R1" & Metric == "viability") %>% pull(Estimate)*100,
           colour = "#ce8e58", size=0.5, alpha=0.5) +
  
  # add a line for SS mean
  geom_hline(yintercept = 89.4, color = "black", linetype = "solid", size = 0.25) +
 
  # This plots individual values for each genotype
  geom_jitter(data = data.viability %>% filter(Genotype %in% c("SS", "SR1", "R1R1")),
             aes(x=Genotype, y = Survival, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +
  

  # This plots the mean viability for each genotype
  geom_point(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR1", "R1R1")),
              aes(x=Genotype.Allele, y = Estimate*100, color = Genotype.Allele, shape = Genotype.Allele), size = 8, alpha = 1) +

  # This plots the confidence interval for each genotype
  geom_errorbar(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR1", "R1R1")),
                aes(x=Genotype.Allele, ymin = Lower.CI*100, ymax = Upper.CI*100,
                    color = Genotype.Allele), width = 0) +


  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
  # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R1
  geom_text(aes(x = 0.7, y = 120, 
              label = "s = -0.15 (-0.19, -0.09)*** \nh = 0.52 (-0.08, 0.60)"), 
          size = 8, hjust = 0, vjust = 1) +
  
  
  scale_y_continuous(breaks = c(40, 50, 60, 70, 80, 90, 100), 
                   limits = c(40, 120)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Viability (%)") +
  
  
  theme_classic() +
  
  theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")
    
  

R1_viability

ggsave("./plots/R1_viability.png",R1_viability, width = 5 , height = 5)
```

### 4.2 Ace R2 genotype

```{r}
genotype_labels = c("SS","SR2","R2R2")

# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#30a4bc", "#ec4176")
genotype_colors = c("#165094", "#165094","#e1615e")

# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)



R2_viability = ggplot() +
  
  annotate("segment", x = 1, xend = 3, 
           y = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "viability") %>% pull(Estimate)*100, 
            yend = all.metrics %>% filter(Genotype.Allele == "R2R2" & Metric == "viability") %>% pull(Estimate)*100,
           colour = "#e1615e", size=0.5, alpha=0.5) +
  
  # add a line for SS mean
  geom_hline(yintercept = 89.4, color = "black", linetype = "solid", size = 0.25) +
 
  # This plots individual values for each genotype
  geom_jitter(data = data.viability %>% filter(Genotype %in% c("SS", "SR2", "R2R2")),
             aes(x=Genotype, y = Survival, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +
  

  # This plots the mean viability for each genotype
  geom_point(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR2", "R2R2")),
              aes(x=Genotype.Allele, y = Estimate*100, color = Genotype.Allele, shape = Genotype.Allele), size = 8, alpha = 1) +

  # This plots the confidence interval for each genotype
  geom_errorbar(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR2", "R2R2")),
                aes(x=Genotype.Allele, ymin = Lower.CI*100, ymax = Upper.CI*100,
                    color = Genotype.Allele), width = 0) +


  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
  # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R2
  geom_text(aes(x = 0.7, y = 120, 
              label = "s = -0.17 (-0.27, -0.09)*** \nh = 0.20 (-0.15, 0.44)*"), 
          size = 8, hjust = 0, vjust = 1) +
  
  
  scale_y_continuous(breaks = c(40, 50, 60, 70, 80, 90, 100), 
                   limits = c(40, 120)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Viability (%)") +
  
  
  theme_classic() +
  
  theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")
    
  

R2_viability

ggsave("./plots/R2_viability.png",R2_viability, width = 5 , height = 5)
```

### 4.3 Ace R3 genotype

```{r}
genotype_labels = c("SS","SR3","R3R3")

# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#97D1DD",  "#782e6e")
genotype_colors = c("#165094", "#165094", "#702771")

# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)



R3_viability = ggplot() +
  
  annotate("segment", x = 0.7, xend = 3, 
           y = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "viability") %>% pull(Estimate)*100, 
            yend = all.metrics %>% filter(Genotype.Allele == "R3R3" & Metric == "viability") %>% pull(Estimate)*100,
           colour = "#782e6e", size=0.5, alpha=0.5) +
  
  # add a line for SS mean
  geom_hline(yintercept = 89.4, color = "black", linetype = "solid", size = 0.25) +
 
  # This plots individual values for each genotype
  geom_jitter(data = data.viability %>% filter(Genotype %in% c("SS", "SR3", "R3R3")),
             aes(x=Genotype, y = Survival, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +
  

  # This plots the mean viability for each genotype
  geom_point(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR3", "R3R3")),
              aes(x=Genotype.Allele, y = Estimate*100, color = Genotype.Allele, shape = Genotype.Allele), size = 8, alpha = 1) +

  # This plots the confidence interval for each genotype
  geom_errorbar(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR3", "R3R3")),
                aes(x=Genotype.Allele, ymin = Lower.CI*100, ymax = Upper.CI*100,
                    color = Genotype.Allele), width = 0) +


  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
  # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R3
  geom_text(aes(x = 0.7, y = 120, 
              label = "s = -0.15 (-0.17, -0.08)*** \nh = 0.50 (-0.12, 0.61)"), 
          size = 8, hjust = 0, vjust = 1) +
  
  
  scale_y_continuous(breaks = c(40, 50, 60, 70, 80, 90, 100), 
                   limits = c(40, 120)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Viability (%)") +
  
  
  theme_classic() +
  
  theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")
    
  

R3_viability

ggsave("./plots/R3_viability.png",R3_viability, width = 5 , height = 5)
```

# 5. References

Regression with categorical variable:

<https://www.youtube.com/watch?v=9yTui_LoSOc>

<http://www.sthda.com/english/articles/40-regression-analysis/163-regression-with-categorical-variables-dummy-coding-essentials-in-r/>
