---
title: "Beneficial Dominance Reversal_Cartoon"
author: "Marianna Karageorgi"
date: "2024-08-31"
output: html_document
---

- If the fitness of the heterozygote is the same as the homozygote sensitive, the resistant allele is recessive.
- If the fitness of the heterozygote is the average of the sensitive and resistant homozygotes, the resistant allele is additive.
- If the fitness of the heterozygote is the same as the homozygote resistant, the resistant allele is dominant.


### Dummy Data 
```{r}
library(tidyverse)
```

```{r}
## R beneficial
data.Rdom = tibble(
  Genotype = factor(c("SS", "SR","SR","SR", "RR"), levels = c("SS", "SR","RR")),
  Dominance = c("recessive", "recessive","additive", "dominant", "dominant"),
  Trait.1 = c(0, 0.1,0.5,0.9, 1)
)

## R deleterious
data.Sdom = tibble(
  Genotype = factor(c("SS", "SR","SR","SR", "RR"), levels = c("SS", "SR","RR")),
  Dominance = c("recessive", "recessive","additive", "dominant", "dominant"),
  Trait.2 = c(1, 0.9,0.5,0.1, 0)
)
```


# Plot
```{r}
library("ggpubr")

genotype_labels = c("SS","SR","RR")



# Define  colors for each dominance category
dominance_colors = c("recessive" = "#165094", "additive" = "#FCC876", "dominant" = "#faa41a")

# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)

# Define text labels for the SR group based on their dominance
labels = c("recessive", "additive", "dominant")


# Plot
Rdom = ggplot() +
  
  annotate("segment", x = 1, xend = 3, y = 0, yend = 1, colour = "#faa41a", size=0.5, alpha=0.5) +
  
  # add a line for SS 
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  
  # This plots data points
  geom_point(data = data.Rdom, 
             aes(x = Genotype, y = Trait.1, color = Dominance, shape = Genotype),
             size = 8, alpha = 1) +
  
  geom_text(data = data.Rdom %>% filter(Genotype == "SR"), 
            aes(x = Genotype, y = Trait.1, label = Dominance, color = Dominance), 
            vjust = 0.3, hjust = 1.2, size = 8, check_overlap = TRUE) +
  
  # Map the colors to dominance
  scale_color_manual(values = dominance_colors) +
  
  # Map the shapes to genotypes
  scale_shape_manual(values = genotype_shapes) +
  
   scale_y_continuous(limits = c(-0.1,1.1)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Phenotypic value") +
  
  theme_classic() +
  
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    
    axis.title.y = element_text(size = 24),
    axis.text.x = element_text(size = 22),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "no",
    plot.title = element_text(hjust = 0.5, size = 18, face = "bold")
  )

Rdom

# Save the plot
ggsave("./plots/Rdom.png",Rdom,  width = 5 , height = 5)
```


```{r}
library("ggpubr")

genotype_labels = c("SS","SR","RR")



# Define  colors for each dominance category
dominance_colors = c("recessive" = "#165094", "additive" =  "#7396BF", "dominant" = "#faa41a")

# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)

# Define text labels for the SR group based on their dominance
labels = c("recessive", "additive", "dominant")


# Plot
Sdom = ggplot() +
  
  annotate("segment", x = 1, xend = 3, y = 1, yend = 0, colour = "#faa41a", size=0.5, alpha=0.5) +
  
  # add a line for SS 
  geom_hline(yintercept = 1 , color = "black", linetype = "solid", size = 0.5) +
  
  # This plots data points
  geom_point(data = data.Sdom, 
             aes(x = Genotype, y = Trait.2, color = Dominance, shape = Genotype),
             size = 8, alpha = 1) +
  
  
  geom_text(data = data.Sdom %>% filter(Genotype == "SR"), 
            aes(x = Genotype, y = Trait.2, label = Dominance, color = Dominance), 
            vjust = 0.3, hjust = 1.2, size = 8, check_overlap = TRUE) +
  
  # Map the colors to dominance
  scale_color_manual(values = dominance_colors) +
  
  # Map the shapes to genotypes
  scale_shape_manual(values = genotype_shapes) +
  
   scale_y_continuous(limits = c(-0.1,1.1)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Phenotypic value") +
  
  theme_classic() +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    
    axis.title.y = element_text(size = 24),
    axis.text.x = element_text(size = 22),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "no",
    plot.title = element_text(hjust = 0.5, size = 18, face = "bold")
  )

Sdom

# Save the plot
ggsave("./plots/Sdom.png",Sdom,  width = 5 , height = 5)
```

