---
title: "Comparison of Fecundity between Ace genotypes in the absence of malathion"
author: "Marianna Karageorgi & Zach Mouza"
date: "2023-09-06"
output: 
pdf_document: default
---

# 1. Data

Experiment was conducted from July 17- 20, 2023.

We introduced five 3-day-old females of each genotype in egg-laying bottles fitted with egg-laying caps filled with Drosophila growth media. Each day over a period of 3 days we counted the number of eggs laid. Each of the genotypes were tested in 8-10 replicate egg-laying bottles. The egg-laying caps were examined under a stereoscope where we counted the eggs laid each day over the 3-day period.

The bottles were randomized and the experimenter, Anastasia Andreeva, was blind to the genotype in each bottle.

### Data wrangling

```{r}
library(tidyverse)
library(data.table)
```

```{r}
fecundity_data = fread("./data/Fecundity_FINAL.csv")

fecundity.data = fecundity_data %>%
  mutate(eggs = rowSums(fecundity_data[, 5:7], na.rm = TRUE), 
         Total = 50 ,
         eggs_laid = eggs / Total * 100 ) %>%
  filter(Nr_flies_Experiment_End == 5) %>%
  dplyr::select(Genotype, eggs_laid) %>%
  slice(-c(60, 63)) %>% # experimental error - wrong nr of flies in egg lay cages
  mutate(Genotype = factor(Genotype, levels = c("SS", "SR1", "SR2", "SR3", "R1R1", "R2R2", "R3R3"))) # Convert Genotype to a factor

```

### Summary Statistics

```{r}
data.fecundity.summary = group_by(fecundity.data, Genotype) %>%
  summarise(
    count = n(),
    mean = mean(eggs_laid, na.rm = TRUE),
    sd = sd(eggs_laid, na.rm = TRUE),
    se = sd / sqrt(8)
  )
```

### Fecundity data: Exploratory Data Analysis

```{r}

# Define the custom colors for the palette
genotype_colors = c("black", "#FCCC1A","#8601AF","#FE2712",  "#FCCC1A", "#8601AF",   "#FE2712")

# Define the linetypes for the boxes
linetypes = c(1, 2, 2, 2, 1, 1, 1)

# Use the chosen colors and linetypes in your plot
fecundity_plot = ggplot(fecundity.data, aes(Genotype, eggs_laid)) +
  geom_boxplot(aes(linetype = Genotype, color = Genotype), fill = NA) +
  geom_point(aes(color = Genotype)) +  # Add a color aesthetic here
  scale_color_manual(values = genotype_colors) +  # Map the colors to the genotypes
  scale_linetype_manual(values = linetypes) +
  ylim(0, 350)+
  labs(x = NULL, y = "Eggs Laid") +
  theme_classic() +
  theme(legend.position = "right") + 
  
  theme(legend.position = "none")
```

# 2. Statistics

Analyze fecundity data for homozygous sensitive SS, homozygous resistant RR, and heterozygous SR genotypes for each resistant allele using linear regression model.

The model is:

eggs=b0​+b1​⋅SR+b2​⋅RR+ϵ

The expected number of eggs for each genotype is represented:

-   For SS (reference group): eggsSS​=b0​

-   For SR: eggsSR=b0+b1

-   For RR: eggsRR=b0+b2

Perform post hoc pairwise t-tests (adjusted following Benjamini-Hochberg) to compare the genotypes.

We also estimate additional parameters using coefficients from the linear model.

**Additivity: Comparison between SR vs. (SS+RR)/2**

Additivity is calculated as the difference between the SR genotype and the average of SS and RR genotypes. It can be compared to the expected value under additivity.

-   Expected value under additivity:

    SR- (SS+RR)/2 = b0+b1 - (b0+(b0+b2))/2 = b0+b1 - (2b0+b2)/2

-   Use bootstrap methods to calculate confidence intervals for the additivity estimate

-   Significance testing with two one-sided t-test: H0 = the estimate for additivity is 0. Ha1 = the estimate for dominance is higher than 0 Ha2 = the estimate for dominance is lower than 0.

**Selection coefficient: Estimate ratio (RR - SS)/ SS**

The selection coefficient is the ratio of the difference between RR and SS genotypes to the SS genotype. This parameter indicates how selection affects the fitness of the resistant genotype compared to the sensitive genotype.

-   Estimate ratio (RR-SS)/SS = ((b0+b2)-b0)/b0=b2/b0

-   Use bootstrap methods to calculate confidence intervals for the selection coefficient

-   Significance testing with two-sided t-test: H0 = the estimate for selection equals to 0. Ha = the estimate for selection is different from 0.

**Dominance coefficient: Estimate (SR−SS)/(RR−SS) (h)**

The dominance coefficient measures the degree to which the heterozygous genotype (SR) exhibits fitness that is intermediate between the homozygous genotypes (SS and RR).

-   Estimate ratio (SR-SS)/(RR-SS)= ((b0+b1)−b0)/((b0+b2)−b0)=b1/b2

-   Use bootstrap methods to calculate confidence intervals for the dominance coefficient

-   Significance testing with two one-sided t-test: H0 = the estimate for dominance is 0.5. Ha1 = the estimate for dominance is higher than 0.5. Ha2 = the estimate for dominance is lower than 0.5

```{r}
library(broom)
library(multcomp) 
library(emmeans)
library(msm)
library(boot)
library(ggpubr)
library(boot.pval)
library(modelsummary)
```

### 2.1 Ace R1 genotype

```{r}
# Data R1
fecundity.data.R1 = fecundity.data %>%
  # Filter data 
  filter(Genotype %in% c("SS", "SR1", "R1R1")) %>%
  # Level data so the SS is the reference
  mutate(Genotype = factor(Genotype, levels = c("SS", "SR1", "R1R1")))

# Fit linear model
model.R1 = lm(eggs_laid ~ Genotype, data = fecundity.data.R1)


# Estimated Marginal Means
emmeans(model.R1, ~ Genotype)

# Extract values for table
emm.R1 = tidy(emmeans(model.R1, ~ Genotype))

SS.mean = round(emm.R1 %>% filter(Genotype == "SS") %>% pull(estimate),2)
SR1.mean = round(emm.R1 %>% filter(Genotype == "SR1") %>% pull(estimate),2)
R1R1.mean = round(emm.R1 %>% filter(Genotype == "R1R1") %>% pull(estimate),2)

SS.se = round(emm.R1 %>% filter(Genotype == "SS") %>% pull(std.error),2)
SR1.se = round(emm.R1 %>% filter(Genotype == "SR1") %>% pull(std.error),2)
R1R1.se = round(emm.R1 %>% filter(Genotype == "R1R1") %>% pull(std.error),2)


# Perform pairwise comparisons
pairwise_comparisons.R1 = pairs(emmeans(model.R1, ~ Genotype))

# Adjusting for multiple comparisons using FDR (Benjamini-Hochberg)
pwc.R1 = summary(pairwise_comparisons.R1, adjust = "BH")

SS.SR1 = round(pwc.R1 %>% filter(contrast == "SS - SR1") %>% pull(estimate),2)
SS.R1R1 = round(pwc.R1 %>% filter(contrast == "SS - R1R1") %>% pull(estimate),2)
SR1.R1R1 = round(pwc.R1 %>% filter(contrast == "SR1 - R1R1") %>% pull(estimate),2)

SS.SR1.se = round(pwc.R1 %>% filter(contrast == "SS - SR1") %>% pull(SE),2)
SS.R1R1.se = round(pwc.R1 %>% filter(contrast == "SS - R1R1") %>% pull(SE),2)
SR1.R1R1.se = round(pwc.R1 %>% filter(contrast == "SR1 - R1R1") %>% pull(SE),2)

SS.SR1.p = round(pwc.R1 %>% filter(contrast == "SS - SR1") %>% pull(p.value),2)
SS.R1R1.p = round(pwc.R1 %>% filter(contrast == "SS - R1R1") %>% pull(p.value),2)
SR1.R1R1.p = round(pwc.R1 %>% filter(contrast == "SR1 - R1R1") %>% pull(p.value),2)
```

**Bootstrapping to calculate confidence intervals**

Define a function that will fit the linear model to a re-sampled dataset and compute the additivity, selection coefficient, and dominance coefficient and confidence intervals.

```{r}
# Define the bootstrap function
bootstrap_function.R1 = function(data, indices) {
  sample_data = data[indices, ]  # Resample data
  model = lm(eggs_laid ~ Genotype, data = sample_data)
  coefficients = coef(model)
  
  # Extract coefficients
  b0 = coefficients["(Intercept)"]
  b1 = coefficients["GenotypeSR1"]
  b2 = coefficients["GenotypeR1R1"]
  
  # Calculate the observed value for SR
  eggs_SR1 = b0 + b1
  
  # Calculate the expected additivity value: (SS + RR) / 2
  expected_additivity = (b0 + b0 + b2) / 2
  
  # Calculate deviation from additivity: SR - (SS + RR) / 2
  deviation_from_additivity = eggs_SR1 - expected_additivity
  
  # Calculate dominance coefficient: (SR - SS) / (RR - SS)
  h = b1 / b2 
  
  # Calculate selection coefficient: (RR - SS) / SS
  s = b2 / b0
  
  return(c(deviation_from_additivity = deviation_from_additivity, selection = s, dominance = h))
}

# Set seed for reproducibility
set.seed(123) 

bootstraps = 1000 

# Perform bootstrap with 1000 replicates
results.R1 = boot(data = fecundity.data.R1, statistic = bootstrap_function.R1, R = bootstraps)
summary(results.R1)

# bootstrap values
deviation_additivity.R1 = as.numeric(results.R1$t0[1])
selection.R1 = as.numeric(results.R1$t0[2])
dominance.R1 = as.numeric(results.R1$t0[3])

# Extract bootstrap replicates (bootstrapped statistics)
boot_estimates = results.R1$t  # This extracts all the bootstrap replicates

# Calculate standard errors for each statistic
se_additivity_R1 = sd(boot_estimates[, 1])  # For additivity
se_selection_R1 = sd(boot_estimates[, 2])   # For selection coefficient
se_dominance_R1 = sd(boot_estimates[, 3])   # For dominance coefficient


# Compute confidence intervals
deviation_additivity_ci.R1 = boot.ci(boot.out = results.R1, type = "perc", index = 1)  # Deviation from additivity
selection_ci.R1 = boot.ci(boot.out = results.R1, type = "perc", index = 2)  # Selection coefficient
dominance_ci.R1 = boot.ci(boot.out = results.R1, type = "perc", index = 3)  # Dominance coefficient

# extract CI values
deviation_additivity_ci.R1.lower_ci = round(deviation_additivity_ci.R1$percent[4],2)  # Lower limit
deviation_additivity_ci.R1.upper_ci = round(deviation_additivity_ci.R1$percent[5],2)  # Upper limit

selection_ci.R1.lower_ci = round(selection_ci.R1$percent[4],2)  # Lower limit
selection_ci.R1.upper_ci = round(selection_ci.R1$percent[5],2)  # Upper limit

dominance_ci.R1.lower_ci = round(dominance_ci.R1$percent[4],2)  # Lower limit
dominance_ci.R1.upper_ci = round(dominance_ci.R1$percent[5],2)  # Upper limit


# Compute p-values for each statistic
pvalue_additivity.R1 = round(boot.pval(results.R1, index = 1, type = "perc", theta_null = 0),4)  # p-value for deviation from additivity
pvalue_selection.R1 = round(boot.pval(results.R1, index = 2, type = "perc", theta_null = 0),4)   # p-value for selection coefficient
pvalue_dominance.R1 = round(boot.pval(results.R1, index = 3, type = "perc", theta_null = 0.5),4)   # p-value for dominance coefficient
```

### 2.2 Ace R2 genotype

```{r}
# Data R2
fecundity.data.R2 = fecundity.data %>%
  # Filter data 
  filter(Genotype %in% c("SS", "SR2", "R2R2")) %>%
  # Level data so the SS is the reference
  mutate(Genotype = factor(Genotype, levels = c("SS", "SR2", "R2R2")))

# Fit linear model
model.R2 = lm(eggs_laid ~ Genotype, data = fecundity.data.R2)

# Estimated Marginal Means
emm.R2 = tidy(emmeans(model.R2, ~ Genotype))

SS.mean = round(emm.R2 %>% filter(Genotype == "SS") %>% pull(estimate), 2)
SR2.mean = round(emm.R2 %>% filter(Genotype == "SR2") %>% pull(estimate), 2)
R2R2.mean = round(emm.R2 %>% filter(Genotype == "R2R2") %>% pull(estimate), 2)

SS.se = round(emm.R2 %>% filter(Genotype == "SS") %>% pull(std.error), 2)
SR2.se = round(emm.R2 %>% filter(Genotype == "SR2") %>% pull(std.error), 2)
R2R2.se = round(emm.R2 %>% filter(Genotype == "R2R2") %>% pull(std.error), 2)

# Perform pairwise comparisons
pairwise_comparisons.R2 = pairs(emmeans(model.R2, ~ Genotype))

# Adjusting for multiple comparisons using FDR (Benjamini-Hochberg)
pwc.R2 = summary(pairwise_comparisons.R2, adjust = "BH")

SS.SR2 = round(pwc.R2 %>% filter(contrast == "SS - SR2") %>% pull(estimate), 2)
SS.R2R2 = round(pwc.R2 %>% filter(contrast == "SS - R2R2") %>% pull(estimate), 2)
SR2.R2R2 = round(pwc.R2 %>% filter(contrast == "SR2 - R2R2") %>% pull(estimate), 2)

SS.SR2.se = round(pwc.R2 %>% filter(contrast == "SS - SR2") %>% pull(SE), 2)
SS.R2R2.se = round(pwc.R2 %>% filter(contrast == "SS - R2R2") %>% pull(SE), 2)
SR2.R2R2.se = round(pwc.R2 %>% filter(contrast == "SR2 - R2R2") %>% pull(SE), 2)

SS.SR2.p = round(pwc.R2 %>% filter(contrast == "SS - SR2") %>% pull(p.value), 2)
SS.R2R2.p = round(pwc.R2 %>% filter(contrast == "SS - R2R2") %>% pull(p.value), 2)
SR2.R2R2.p = round(pwc.R2 %>% filter(contrast == "SR2 - R2R2") %>% pull(p.value), 2)
```

**Bootstrapping to calculate confidence intervals**

Define a function that will fit the linear model to a resampled dataset and compute the additivity, selection coefficient, and dominance coefficient and confidence intervals.

```{r}
# Define the bootstrap function
bootstrap_function.R2 = function(data, indices) {
  sample_data = data[indices, ]  # Resample data
  model = lm(eggs_laid ~ Genotype, data = sample_data)
  coefficients = coef(model)
  
  # Extract coefficients
  b0 = coefficients["(Intercept)"]
  b1 = coefficients["GenotypeSR2"]
  b2 = coefficients["GenotypeR2R2"]
  
  # Calculate the observed value for SR
  eggs_SR2 = b0 + b1
  
  # Calculate the expected additivity value: (SS + RR) / 2
  expected_additivity = (b0 + b0 + b2) / 2
  
  # Calculate deviation from additivity: SR - (SS + RR) / 2
  deviation_from_additivity = eggs_SR2 - expected_additivity
  
  # Calculate dominance coefficient: (SR - SS) / (RR - SS)
  h = b1 / b2 
  
  # Calculate selection coefficient: (RR - SS) / SS
  s = b2 / b0
  
  return(c(deviation_from_additivity = deviation_from_additivity, selection = s, dominance = h))
}

# Set seed for reproducibility
set.seed(123) 

bootstraps = 1000 

# Perform bootstrap with 1000 replicates
results.R2 = boot(data = fecundity.data.R2, statistic = bootstrap_function.R2, R = bootstraps)
summary(results.R2)

# Bootstrap values
deviation_additivity.R2 = as.numeric(results.R2$t0[1])
selection.R2 = as.numeric(results.R2$t0[2])
dominance.R2 = as.numeric(results.R2$t0[3])

# Extract bootstrap replicates (bootstrapped statistics)
boot_estimates = results.R2$t  # This extracts all the bootstrap replicates

# Calculate standard errors for each statistic
se_additivity_R2 = sd(boot_estimates[, 1])  # For additivity
se_selection_R2 = sd(boot_estimates[, 2])   # For selection coefficient
se_dominance_R2 = sd(boot_estimates[, 3])   # For dominance coefficient

# Compute confidence intervals
deviation_additivity_ci.R2 = boot.ci(boot.out = results.R2, type = "perc", index = 1)  # Deviation from additivity
selection_ci.R2 = boot.ci(boot.out = results.R2, type = "perc", index = 2)  # Selection coefficient
dominance_ci.R2 = boot.ci(boot.out = results.R2, type = "perc", index = 3)  # Dominance coefficient

# Extract CI values
deviation_additivity_ci.R2.lower_ci = round(deviation_additivity_ci.R2$percent[4], 2)  # Lower limit
deviation_additivity_ci.R2.upper_ci = round(deviation_additivity_ci.R2$percent[5], 2)  # Upper limit

selection_ci.R2.lower_ci = round(selection_ci.R2$percent[4], 2)  # Lower limit
selection_ci.R2.upper_ci = round(selection_ci.R2$percent[5], 2)  # Upper limit

dominance_ci.R2.lower_ci = round(dominance_ci.R2$percent[4], 2)  # Lower limit
dominance_ci.R2.upper_ci = round(dominance_ci.R2$percent[5], 2)  # Upper limit

# Compute p-values for each statistic
pvalue_additivity.R2 = round(boot.pval(results.R2, index = 1, type = "perc", theta_null = 0), 4)  # p-value for deviation from additivity
pvalue_selection.R2 = round(boot.pval(results.R2, index = 2, type = "perc", theta_null = 0), 4)   # p-value for selection coefficient
pvalue_dominance.R2 = round(boot.pval(results.R2, index = 3, type = "perc", theta_null = 0.5), 4)   # p-value for dominance coefficient
```

### 2.3 Ace R3 genotype

```{r}
# Data R3
fecundity.data.R3 = fecundity.data %>%
  # Filter data 
  filter(Genotype %in% c("SS", "SR3", "R3R3")) %>%
  # Level data so the SS is the reference
  mutate(Genotype = factor(Genotype, levels = c("SS", "SR3", "R3R3")))

# Fit linear model
model.R3 = lm(eggs_laid ~ Genotype, data = fecundity.data.R3)

# Estimated Marginal Means
emm.R3 = tidy(emmeans(model.R3, ~ Genotype))

SS.mean = round(emm.R3 %>% filter(Genotype == "SS") %>% pull(estimate), 2)
SR3.mean = round(emm.R3 %>% filter(Genotype == "SR3") %>% pull(estimate), 2)
R3R3.mean = round(emm.R3 %>% filter(Genotype == "R3R3") %>% pull(estimate), 2)

SS.se = round(emm.R3 %>% filter(Genotype == "SS") %>% pull(std.error), 2)
SR3.se = round(emm.R3 %>% filter(Genotype == "SR3") %>% pull(std.error), 2)
R3R3.se = round(emm.R3 %>% filter(Genotype == "R3R3") %>% pull(std.error), 2)

# Perform pairwise comparisons
pairwise_comparisons.R3 = pairs(emmeans(model.R3, ~ Genotype))

# Adjusting for multiple comparisons using FDR (Benjamini-Hochberg)
pwc.R3 = summary(pairwise_comparisons.R3, adjust = "BH")

SS.SR3 = round(pwc.R3 %>% filter(contrast == "SS - SR3") %>% pull(estimate), 2)
SS.R3R3 = round(pwc.R3 %>% filter(contrast == "SS - R3R3") %>% pull(estimate), 2)
SR3.R3R3 = round(pwc.R3 %>% filter(contrast == "SR3 - R3R3") %>% pull(estimate), 2)

SS.SR3.se = round(pwc.R3 %>% filter(contrast == "SS - SR3") %>% pull(SE), 2)
SS.R3R3.se = round(pwc.R3 %>% filter(contrast == "SS - R3R3") %>% pull(SE), 2)
SR3.R3R3.se = round(pwc.R3 %>% filter(contrast == "SR3 - R3R3") %>% pull(SE), 2)

SS.SR3.p = round(pwc.R3 %>% filter(contrast == "SS - SR3") %>% pull(p.value), 2)
SS.R3R3.p = round(pwc.R3 %>% filter(contrast == "SS - R3R3") %>% pull(p.value), 2)
SR3.R3R3.p = round(pwc.R3 %>% filter(contrast == "SR3 - R3R3") %>% pull(p.value), 2)
```

**Bootstrapping to calculate confidence intervals**

Define a function that will fit the linear model to a resampled dataset and compute the additivity, selection coefficient, and dominance coefficient and confidence intervals.

```{r}
# Define the bootstrap function
bootstrap_function.R3 = function(data, indices) {
  sample_data = data[indices, ]  # Resample data
  model = lm(eggs_laid ~ Genotype, data = sample_data)
  coefficients = coef(model)
  
  # Extract coefficients
  b0 = coefficients["(Intercept)"]
  b1 = coefficients["GenotypeSR3"]
  b2 = coefficients["GenotypeR3R3"]
  
  # Calculate the observed value for SR
  eggs_SR3 = b0 + b1
  
  # Calculate the expected additivity value: (SS + RR) / 2
  expected_additivity = (b0 + b0 + b2) / 2
  
  # Calculate deviation from additivity: SR - (SS + RR) / 2
  deviation_from_additivity = eggs_SR3 - expected_additivity
  
  # Calculate dominance coefficient: (SR - SS) / (RR - SS)
  h = b1 / b2 
  
  # Calculate selection coefficient: (RR - SS) / SS
  s = b2 / b0
  
  return(c(deviation_from_additivity = deviation_from_additivity, selection = s, dominance = h))
}

# Set seed for reproducibility
set.seed(123) 

bootstraps = 1000 

# Perform bootstrap with 1000 replicates
results.R3 = boot(data = fecundity.data.R3, statistic = bootstrap_function.R3, R = bootstraps)
summary(results.R3)

# Bootstrap values
deviation_additivity.R3 = as.numeric(results.R3$t0[1])
selection.R3 = as.numeric(results.R3$t0[2])
dominance.R3 = as.numeric(results.R3$t0[3])

# Extract bootstrap replicates (bootstrapped statistics)
boot_estimates = results.R3$t  # This extracts all the bootstrap replicates

# Calculate standard errors for each statistic
se_additivity_R3 = sd(boot_estimates[, 1])  # For additivity
se_selection_R3 = sd(boot_estimates[, 2])   # For selection coefficient
se_dominance_R3 = sd(boot_estimates[, 3])   # For dominance coefficient

# Compute confidence intervals
deviation_additivity_ci.R3 = boot.ci(boot.out = results.R3, type = "perc", index = 1)  # Deviation from additivity
selection_ci.R3 = boot.ci(boot.out = results.R3, type = "perc", index = 2)  # Selection coefficient
dominance_ci.R3 = boot.ci(boot.out = results.R3, type = "perc", index = 3)  # Dominance coefficient

# Extract CI values
deviation_additivity_ci.R3.lower_ci = round(deviation_additivity_ci.R3$percent[4], 2)  # Lower limit
deviation_additivity_ci.R3.upper_ci = round(deviation_additivity_ci.R3$percent[5], 2)  # Upper limit

selection_ci.R3.lower_ci = round(selection_ci.R3$percent[4], 2)  # Lower limit
selection_ci.R3.upper_ci = round(selection_ci.R3$percent[5], 2)  # Upper limit

dominance_ci.R3.lower_ci = round(dominance_ci.R3$percent[4], 2)  # Lower limit
dominance_ci.R3.upper_ci = round(dominance_ci.R3$percent[5], 2)  # Upper limit

# Compute p-values for each statistic
pvalue_additivity.R3 = round(boot.pval(results.R3, index = 1, type = "perc", theta_null = 0), 4)  # p-value for deviation from additivity
pvalue_selection.R3 = round(boot.pval(results.R3, index = 2, type = "perc", theta_null = 0), 4)   # p-value for selection coefficient
pvalue_dominance.R3 = round(boot.pval(results.R3, index = 3, type = "perc", theta_null = 0.5), 4)   # p-value for dominance coefficient
```

# 3. Tables with statistics

```{r}
library(kableExtra)
library(htmltools)
library(webshot)
source("./StatTable.R") # table for statistics
```

## 3.1 Ace R1 genotype

```{r}
# Call StatTable for model.R1
StatTable(model = model.R1, 
          directory = "./tables/", 
          display_title = "Effect of genotype on fecundity",  # Title for the table
          save_title = "Fecundity.R1",  # Title for the saved file
          pred_labels = c("Intercept","SR1", "R1R1"), # replace with your labels
          dv_labels = "Genotype")

# dataframe with estimates
table.R1 = data.frame(
  Parameter = c("SS mean", "SR1 mean", "R1R1 mean", 
                "SS vs. SR1", "SS vs. R1R1", "SR1 vs. R1R1", 
                "SR1 - (SS + R1R1)/2 (additivity)", 
                "(R1R1 - SS) / R1R1 (selection, s)", "(SR1 - SS) / (R1R1 - SS) (dominance, h)"), 
  Fecundity = c(SS.mean, SR1.mean, R1R1.mean, 
                SS.SR1, SS.R1R1, SR1.R1R1, 
                deviation_additivity.R1, selection.R1, dominance.R1), 
  Std.Error = c(SS.se, SR1.se, R1R1.se, 
                SS.SR1.se, SS.R1R1.se, SR1.R1R1.se, 
                se_additivity_R1, se_selection_R1, se_dominance_R1),
  Lower.CI = c("-", "-", "-", "-", "-", "-",
                deviation_additivity_ci.R1.lower_ci,
                selection_ci.R1.lower_ci,
                dominance_ci.R1.lower_ci),
  Upper.CI = c("-", "-", "-", "-", "-", "-",
                 deviation_additivity_ci.R1.upper_ci,
                 selection_ci.R1.upper_ci,
                 dominance_ci.R1.upper_ci),
  P.value = c("-", "-", "-", SS.SR1.p, SS.R1R1.p, SR1.R1R1.p, 
         pvalue_additivity.R1, pvalue_selection.R1, pvalue_dominance.R1)
)

# Format the columns
table.R1$Fecundity = formatC(table.R1$Fecundity, format = "f", digits = 2)
table.R1$Std.Error = formatC(table.R1$Std.Error, format = "f", digits = 2)
table.R1$Lower.CI = formatC(table.R1$Lower.CI, format = "f", digits = 2)
table.R1$Upper.CI = formatC(table.R1$Upper.CI, format = "f", digits = 2)
table.R1$P.value = formatC(table.R1$P.value, format = "e", digits = 4)  # 4 decimals for p-values


table.R1.kable = kable(x = table.R1, digits = 4, align = 'l',
                             caption = "<b>Effect estimates from linear model analysis for fecundity</b>") %>%
                       kable_classic(full_width = TRUE, html_font = "Times New Roman", font_size = 16) %>%
                       kable_styling(latex_options = c("striped", "hold_position"), position = "center") %>%
                       row_spec(0, bold = TRUE, font_size = 16) %>%  # Make header bold and set font size
                       row_spec(1:nrow(table.R1), font_size = 16)  # Smaller font for values
table.R1.kable
```

## 3.2 Ace R2 genotype

```{r}
# Call StatTable for model.R2
StatTable(model = model.R2, 
          directory = "./tables/", 
          display_title = "Effect of genotype on fecundity",  # Title for the table
          save_title = "Fecundity.R2",  # Title for the saved file
          pred_labels = c("Intercept", "SR2", "R2R2"), # Updated labels
          dv_labels = "Genotype")

# Dataframe with estimates
table.R2 = data.frame(
  Parameter = c("SS mean", "SR2 mean", "R2R2 mean", 
                "SS vs. SR2", "SS vs. R2R2", "SR2 vs. R2R2", 
                "SR2 - (SS + R2R2)/2 (additivity)", 
                "(R2R2 - SS) / R2R2 (selection, s)", "(SR2 - SS) / (R2R2 - SS) (dominance, h)"), 
  Fecundity = c(SS.mean, SR2.mean, R2R2.mean, 
                SS.SR2, SS.R2R2, SR2.R2R2, 
                deviation_additivity.R2, selection.R2, dominance.R2), 
  Std.Error = c(SS.se, SR2.se, R2R2.se, 
                SS.SR2.se, SS.R2R2.se, SR2.R2R2.se, 
                se_additivity_R2, se_selection_R2, se_dominance_R2),
  Lower.CI = c("-", "-", "-", "-", "-", "-",
                deviation_additivity_ci.R2.lower_ci,
                selection_ci.R2.lower_ci,
                dominance_ci.R2.lower_ci),
  Upper.CI = c("-", "-", "-", "-", "-", "-",
                 deviation_additivity_ci.R2.upper_ci,
                 selection_ci.R2.upper_ci,
                 dominance_ci.R2.upper_ci),
  P.value = c("-", "-", "-", SS.SR2.p, SS.R2R2.p, SR2.R2R2.p, 
               pvalue_additivity.R2, pvalue_selection.R2, pvalue_dominance.R2)
)

# Format the columns
table.R2$Fecundity = formatC(table.R2$Fecundity, format = "f", digits = 2)
table.R2$Std.Error = formatC(table.R2$Std.Error, format = "f", digits = 2)
table.R2$Lower.CI = formatC(table.R2$Lower.CI, format = "f", digits = 2)
table.R2$Upper.CI = formatC(table.R2$Upper.CI, format = "f", digits = 2)
table.R2$P.value = formatC(table.R2$P.value, format = "e", digits = 4)  # 4 decimals for p-values

# Use kable to create the table
table.R2.kable = kable(x = table.R2, digits = 4, align = 'l',
                 caption = "<b>Effect estimates from linear model analysis for fecundity</b>") %>%
            kable_classic(full_width = TRUE, html_font = "Times New Roman", font_size = 16) %>%
            kable_styling(latex_options = c("striped", "hold_position"), position = "center") %>%
            row_spec(0, bold = TRUE, font_size = 16) %>%  # Make header bold and set font size
            row_spec(1:nrow(table.R2), font_size = 16)  # Smaller font for values
```

## 3.3 Ace R3 genotype

```{r}
# Call StatTable for model.R3
StatTable(model = model.R3, 
          directory = "./tables/", 
          display_title = "Effect of genotype on fecundity",  # Title for the table
          save_title = "Fecundity.R3",  # Title for the saved file
          pred_labels = c("Intercept", "SR3", "R3R3"), # Updated labels
          dv_labels = "Genotype")

# Dataframe with estimates
table.R3 = data.frame(
  Parameter = c("SS mean", "SR3 mean", "R3R3 mean", 
                "SS vs. SR3", "SS vs. R3R3", "SR3 vs. R3R3", 
                "SR3 - (SS + R3R3)/2 (additivity)", 
                "(R3R3 - SS) / R3R3 (selection, s)", "(SR3 - SS) / (R3R3 - SS) (dominance, h)"), 
  Fecundity = c(SS.mean, SR3.mean, R3R3.mean, 
                SS.SR3, SS.R3R3, SR3.R3R3, 
                deviation_additivity.R3, selection.R3, dominance.R3), 
  Std.Error = c(SS.se, SR3.se, R3R3.se, 
                SS.SR3.se, SS.R3R3.se, SR3.R3R3.se, 
                se_additivity_R3, se_selection_R3, se_dominance_R3),
  Lower.CI = c("-", "-", "-", "-", "-", "-",
                deviation_additivity_ci.R3.lower_ci,
                selection_ci.R3.lower_ci,
                dominance_ci.R3.lower_ci),
  Upper.CI = c("-", "-", "-", "-", "-", "-",
                 deviation_additivity_ci.R3.upper_ci,
                 selection_ci.R3.upper_ci,
                 dominance_ci.R3.upper_ci),
  P.value = c("-", "-", "-", SS.SR3.p, "<.0001", "<.0001", 
               "1e-16", pvalue_selection.R3, pvalue_dominance.R3)
)

# Format the columns
table.R3$Fecundity = formatC(table.R3$Fecundity, format = "f", digits = 2)
table.R3$Std.Error = formatC(table.R3$Std.Error, format = "f", digits = 2)
table.R3$Lower.CI = formatC(table.R3$Lower.CI, format = "f", digits = 2)
table.R3$Upper.CI = formatC(table.R3$Upper.CI, format = "f", digits = 2)
table.R3$P.value = formatC(table.R3$P.value, format = "e", digits = 4)  # 4 decimals for p-values

# Use kable to create the table
table.R3.kable = kable(x = table.R3, digits = 4, align = 'l',
                 caption = "<b>Effect estimates from linear model analysis for fecundity</b>") %>%
            kable_classic(full_width = TRUE, html_font = "Times New Roman", font_size = 16) %>%
            kable_styling(latex_options = c("striped", "hold_position"), position = "center") %>%
            row_spec(0, bold = TRUE, font_size = 16) %>%  # Make header bold and set font size
            row_spec(1:nrow(table.R3), font_size = 16)  # Smaller font for values
```

## 3.4 R1+R2+R3 Table

```{r}
# remove first row in table 2 and table 3 because it is the same with table 1
table.R1.R2.R3 = rbind(table.R1, table.R2 %>% slice(-1), table.R3 %>% slice(-1))

# Use kable to create the table
table.R1.R2.R3.kable = table.R1.R2.R3 %>%
  kbl(caption = "Effect estimates from linear model analysis for fecundity") %>%
  kable_classic(full_width = TRUE, html_font = "Arial") 
   

table.R1.R2.R3.kable


table.R1.R2.R3.kable

# Save the table as an HTML file
save_kable(table.R1.R2.R3.kable, file = "./tables/table_fecundity.html")

# Use webshot to convert HTML to PDF
webshot("./tables/table_fecundity.html", file = "./tables/table_fecundity.pdf")
```

## 3.5 Data for simulations

```{r}
table.fecundity = rbind (table.R1, table.R2, table.R3)
```

# 4. Plots

### 4.1 Ace R1 genotype

```{r}

genotype_labels = c("SS","SR1","R1R1")

# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#97D1DD", "#ffce1b")
genotype_colors = c("#165094", "#7396BF","#ce8e58")


# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)


R1_fecundity = ggplot() +
  
  annotate("segment", x = 1, xend = 3, y = 200.2222, yend = 136.5000, colour = "#ce8e58", size=0.5, alpha=0.5) +
  
  # This plots individual values for each genotype
  geom_jitter(data = fecundity.data %>% filter(Genotype %in% c("SS", "SR1", "R1R1")),
             aes(x=Genotype, y = eggs_laid, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +

  # add a line for SS mean
  geom_hline(yintercept = 200.2222, color = "black", linetype = "solid", size = 0.25) +
  
  # This plots the mean fecundity for each genotype
  geom_point(data = data.fecundity.summary %>% filter(Genotype %in% c("SS","SR1", "R1R1")),
              aes(x=Genotype, y = mean, color = Genotype, shape = Genotype), size = 8, alpha = 1) +
  
  # This plots the standard error for each genotype
  geom_errorbar(data = data.fecundity.summary %>% filter(Genotype %in% c("SS", "SR1", "R1R1")),
                aes(x=Genotype, ymin = mean - se, ymax = mean + se, color = Genotype), width = 0) +
  
  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
   # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R1
  geom_text(aes(x = 1.2, y = 320, 
                label = "s = -0.32 +/-0.09** \nh = 0.55 +/-0.40"), 
            size = 8, hjust = 0, vjust = 1) +

  scale_y_continuous(breaks = seq(50, 320, 50), 
                     limits = c(50,320)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Fecundity") +
  
  theme_classic() +
  
   theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")

R1_fecundity

ggsave("./plots/R1_fecundity.png",R1_fecundity,  width = 5 , height = 5)
```

### 4.2 Ace R2 genotype

```{r}

genotype_labels = c("SS","SR2","R2R2")



# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#97D1DD", "#ec4176")
genotype_colors = c("#165094", "#7396BF","#e1615e")


# Define the shapes for each genotype

genotype_shapes = c(16, 15, 17)


R2_fecundity = ggplot() +
  
  annotate("segment", x = 1, xend = 3, y = 200.2222, yend = 167.7778, colour = "#ec4176", size=0.5, alpha=0.5) +
  
  # This plots individual values for each genotype
  geom_jitter(data = fecundity.data %>% filter(Genotype %in% c("SS", "SR2", "R2R2")),
             aes(x=Genotype, y = eggs_laid, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +

  # add a line for SS mean
  geom_hline(yintercept = 200.2222, color = "black", linetype = "solid", size = 0.25) +
  
  # This plots the mean fecundity for each genotype
  geom_point(data = data.fecundity.summary %>% filter(Genotype %in% c("SS","SR2", "R2R2")),
              aes(x=Genotype, y = mean, color = Genotype, shape = Genotype), size = 8, alpha = 1) +
  
  # This plots the standard error for each genotype
  geom_errorbar(data = data.fecundity.summary %>% filter(Genotype %in% c("SS", "SR2", "R2R2")),
                aes(x=Genotype, ymin = mean - se, ymax = mean + se, color = Genotype), width = 0) +
 
  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
   # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R2
  geom_text(aes(x = 1.2, y = 320, 
                label = "s = -0.16 +/-0.09 \nh = 0.39 +/-0.40"), 
            size = 8, hjust = 0, vjust = 1) +
  
  scale_y_continuous(breaks = seq(50, 320, 50), 
                     limits = c(50,320)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Fecundity") +
  
  theme_classic() +
  
  theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")

R2_fecundity

ggsave("./plots/R2_fecundity.png",R2_fecundity ,width = 5 , height = 5)
```

### 4.3 Ace R3 genotype

```{r}
library("ggpubr")

genotype_labels = c("SS","SR3","R3R3")

# Define the custom colors for each genotype
#genotype_colors = c("black", "#4d4d4d",  "#FE2712")
#genotype_colors = c("#30a4bc", "#30a4bc",  "#782e6e")
genotype_colors = c("#165094", "#165094", "#702771")


# Define the shapes for each genotype

genotype_shapes = c(16, 15, 17)


R3_fecundity = ggplot() +
  
  annotate("segment", x = 1, xend = 3, y = 200.2222, yend = 106.4444, colour = "#782e6e", size=0.5, alpha=0.5) +
  
  # This plots individual values for each genotype
  geom_jitter(data = fecundity.data %>% filter(Genotype %in% c("SS", "SR3", "R3R3")),
             aes(x=Genotype, y = eggs_laid, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +

  # add a line for SS mean
  geom_hline(yintercept = 200.2222, color = "black", linetype = "solid", size = 0.25) +
  
  # This plots the mean fecundity for each genotype
  geom_point(data = data.fecundity.summary %>% filter(Genotype %in% c("SS","SR3", "R3R3")),
              aes(x=Genotype, y = mean, color = Genotype, shape = Genotype), size = 8, alpha = 1) +
  
  # This plots the standard error for each genotype
  geom_errorbar(data = data.fecundity.summary %>% filter(Genotype %in% c("SS", "SR3", "R3R3")),
                aes(x=Genotype, ymin = mean - se, ymax = mean + se, color = Genotype), width = 0) +
  

  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
   # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R3
  geom_text(aes(x = 1.2, y = 320, 
                label = "s = -0.47 +/-0.09*** \nh = -0.02 +/-0.40***"), 
            size = 8, hjust = 0, vjust = 1) +
  
  scale_y_continuous(breaks = seq(50, 320, 50), 
                     limits = c(50,320)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Fecundity") +
  
  theme_classic() +
   
  theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")

R3_fecundity

ggsave("./plots/R3_fecundity.png",R3_fecundity, width = 5 , height = 5)
```

## 5. References

1.  Regression with categorical variable:

    <https://www.youtube.com/watch?v=9yTui_LoSOc>

    <http://www.sthda.com/english/articles/40-regression-analysis/163-regression-with-categorical-variables-dummy-coding-essentials-in-r/>

2.  Posthoc for linear regression with categorical variables:

    <https://biostats.w.uib.no/post-hoc-tests-multiple-comparisons-in-linear-models/?fbclid=IwZXh0bgNhZW0CMTEAAR0AmUvGICwwPPPMhMJPpMzpfRu8T50znT3I2iQFf9uCVBr6VImJjIuQ4G4_aem_jqDm-OcZa61xJDYWr5584Q>

3.  Add p values manually:

    <http://rpkgs.datanovia.com/ggpubr/reference/stat_pvalue_manual.html#examples>

4.  Marginal means:

    <https://easystats.github.io/modelbased/articles/estimate_means.html>

    <https://www.rdocumentation.org/packages/msm/versions/1.7.1/topics/deltamethod>

5.  Bootstrap:

    <https://stats.oarc.ucla.edu/r/faq/how-can-i-generate-bootstrap-statistics-in-r/>

    <https://advstats.psychstat.org/book/ci/bootstrap.php>
