---
title: "Comparison of Egg-adult viability between Ace genotypes in the absence of malathion"
author: "Marianna Karageorgi & Zach Mouza"
date: "2023-08-09"
output:
  pdf_document: default
---

# 1. Data

Starting July 6th 2023, eggs of each D.melanogaster genotype (SS, R1R1, R2R2, R3R3, SR1, SR2, SR3) were exposed to regular Drosophila media (50 eggs/trial). For each genotype we had 10 replicates.

The resistant alleles are Ace-S: IGFG, Ace-R1: [V]{.underline}GFG, Ace-R2: [V]{.underline}GF[A]{.underline}, Ace-R3: [VAY]{.underline}G

### Data Wrangling

```{r}
library(tidyverse)
library(data.table)
```

```{r}
data = fread("./data/Viability_FINAL.csv")

data.viability = data %>%
  mutate(
    Emerge = rowSums(select(., 4:12), na.rm = TRUE),
    Total = 50,
    Not_Emerge = Total - Emerge,
    Survival = Emerge / Total * 100) %>%
  
    mutate(Genotype = factor(Genotype, 
                             levels = c("SS","SR1","SR2","SR3","R1R1", "R2R2", "R3R3"))) %>%
  select(Genotype, Emerge, Not_Emerge, Survival, Total)
```

### Summary Statistics

```{r}
data.viability.summary = group_by(data.viability, Genotype) %>%
  summarise(
    count = n(),
    mean = mean(Survival, na.rm = TRUE),
    sd = sd(Survival, na.rm = TRUE),
    se = sd / sqrt(10)
  )
```

### Viability data: Exploratory Data Analysis

```{r}
library("ggpubr")
library("ggplot2")


# Define the custom colors for the palette
genotype_colors = c("black", "#FCCC1A","#8601AF","#FE2712",  "#FCCC1A", "#8601AF",   "#FE2712")

# Define the linetypes for the boxes
linetypes = c(1, 2, 2, 2, 1, 1, 1)


viability_plot = ggplot(data.viability, 
                        aes(Genotype, Survival)) +
  
  geom_boxplot(aes(linetype = Genotype, color = Genotype), fill = NA) +
  
  geom_point(aes(color = Genotype)) +  
  scale_color_manual(values = genotype_colors) +
  scale_linetype_manual(values = linetypes) +
  labs(x = NULL, y = "Egg to adult vialibity (%)") +
  ylim(40, 100) + 
  theme_classic() 

viability_plot 
```

# 2. Statistics

Analyze viability data for homozygous sensitive SS, homozygous resistant RR, and heterozygous SR genotypes for each resistant allele using logistic regression model. Use a logistic model because of grouped binary data (nr of egg emerged out of 50 eggs).

The model is:

logit(p)=log(p/(1-p)= b0​+b1​⋅SR+b2​⋅RR+ϵ

The odds of emergence for each genotype is represented:

For SS (reference group): odds.emerge.SS​=b0​

For SR: odds.emerge.SR=b0+b1

For RR: odds.emerge.RR=b0+b2

Transform logit values to probabilities of emergence:

proportion_SS = exp(logit_SS) / (1 + exp(logit_SS))

proportion_SR = exp(logit_SR) / (1 + exp(logit_SR))

proportion_RR = exp(logit_RR) / (1 + exp(logit_RR))

Perform post hoc pairwise t-tests (adjusted following Benjamini-Hochberg) to compare the genotypes.

**Additivity: Comparison between SR vs. (SS+RR)/2**

Additivity is calculated as the difference between the SR genotype and the average of SS and RR genotypes. It can be compared to the expected value under additivity.

-   Expected value under additivity:

    SR- (SS+RR)/2 = proportion_SR -(proportion_SS+proportion_RR)/2

-   Use bootstrap methods to calculate confidence intervals for the additivity estimate, and determine whether the confidence interval includes 0.

-   Significance testing with two one-sided t-test: H0 = the estimate for additivity is 0. Ha1 = the estimate for dominance is higher than 0 Ha2 = the estimate for dominance is lower than 0.

**Selection coefficient: Estimate ratio (RR - SS)/ SS**

The selection coefficient is the ratio of the difference between RR and SS genotypes to the SS genotype. This parameter indicates how selection affects the fitness of the resistant genotype compared to the sensitive genotype.

-   Estimate ratio (RR-SS)/SS = (proportion_RR-proportion_SS)/proportion_SS

-   Use bootstrap methods to calculate confidence intervals for the selection coefficient, and determine whether the confidence interval includes 0.

-   Significance testing with two-sided t-test: H0 = the estimate for selection equals to 0. Ha = the estimate for selection is different from 0.

**Dominance coefficient: Estimate (SR−SS)/(RR−SS) (h)**

The dominance coefficient measures the degree to which the heterozygous genotype (SR) exhibits fitness that is intermediate between the homozygous genotypes (SS and RR).

-   Estimate ratio (proportion_SR-proportion_SS)/(proportion_RR-proportion_SS)

-   Use bootstrap methods to calculate confidence intervals for the dominance coefficient, and determine whether the confidence interval includes 0.5.

-   Significance testing with two one-sided t-test: H0 = the estimate for dominance is 0.5. Ha1 = the estimate for dominance is higher than 0.5. Ha2 = the estimate for dominance is lower than 0.5

```{r}
library(broom)
library(multcomp) 
library(emmeans)
library(msm)
library(boot)
library(ggpubr)
library(boot.pval)
library(modelsummary)
```

### 2.1 Ace R1 genotype

```{r}
# Data R1 
data.viability.R1 = data.viability %>% 
  # Filter data 
  filter(Genotype %in% c("SS", "SR1", "R1R1")) %>%
  # Level data so the SS is the ref
  mutate(Genotype = factor(Genotype, levels = c("SS","SR1", "R1R1")))
  
# Fit logistic binomial model
model.R1 = glm(cbind(Emerge, Not_Emerge) ~ Genotype, data = data.viability.R1, family = "binomial")
  
# Calculate estimated marginal means for the Genotype on the response scale (probability) - otherwise the model gives logit values
emmeans(model.R1, ~ Genotype, type = "response")

# Extract values for table
emm.R1 = tidy(emmeans(model.R1, ~ Genotype, type = "response"))

SS.mean = round(emm.R1 %>% filter(Genotype == "SS") %>% pull(prob),2)
SR1.mean = round(emm.R1 %>% filter(Genotype == "SR1") %>% pull(prob),2)
R1R1.mean = round(emm.R1 %>% filter(Genotype == "R1R1") %>% pull(prob),2)

SS.se = round(emm.R1 %>% filter(Genotype == "SS") %>% pull(std.error),2)
SR1.se = round(emm.R1 %>% filter(Genotype == "SR1") %>% pull(std.error),2)
R1R1.se = round(emm.R1 %>% filter(Genotype == "R1R1") %>% pull(std.error),2)

# Perform pairwise comparisons
pairwise_comparisons.R1 = pairs(emmeans(model.R1, ~ Genotype, type = "response"))
  
# Adjusting for multiple comparisons using FDR (Benjamini-Hochberg)
pwc.R1 = summary(pairwise_comparisons.R1, adjust = "BH")
  
SS.SR1 = round(pwc.R1 %>% filter(contrast == "SS / SR1") %>% pull(odds.ratio),2)
SS.R1R1 = round(pwc.R1 %>% filter(contrast == "SS / R1R1") %>% pull(odds.ratio),2)
SR1.R1R1 = round(pwc.R1 %>% filter(contrast == "SR1 / R1R1") %>% pull(odds.ratio),2)

SS.SR1.se = round(pwc.R1 %>% filter(contrast == "SS / SR1") %>% pull(SE),2)
SS.R1R1.se = round(pwc.R1 %>% filter(contrast == "SS / R1R1") %>% pull(SE),2)
SR1.R1R1.se = round(pwc.R1 %>% filter(contrast == "SR1 / R1R1") %>% pull(SE),2)

SS.SR1.p = round(pwc.R1 %>% filter(contrast == "SS / SR1") %>% pull(p.value),4)
SS.R1R1.p = round(pwc.R1 %>% filter(contrast == "SS / R1R1") %>% pull(p.value),4)
SR1.R1R1.p = round(pwc.R1 %>% filter(contrast == "SR1 / R1R1") %>% pull(p.value),4)
```

**Bootstrapping to calculate confidence intervals**

Define a function that will fit the linear model to a resampled dataset and compute the additivity, selection coefficient, and dominance coefficient and confidence intervals.

```{r}
bootstrap_function.R1 = function(data, indices) {
  sample_data = data[indices, ]  # Resample data
  model = glm(cbind(Emerge, Not_Emerge) ~ Genotype, data = sample_data, family = "binomial")
  coefficients = coef(model)
  
  b0 = coefficients["(Intercept)"]
  b1 = coefficients["GenotypeSR1"]
  b2 = coefficients["GenotypeR1R1"]
  
  # Transform logit values to probabilities
  proportion_SS = exp(b0) / (1 + exp(b0))
  proportion_SR1 = exp(b0 + b1) / (1 + exp(b0 + b1))
  proportion_R1R1 = exp(b0 + b2) / (1 + exp(b0 + b2))
  
  # Calculate additivity
  deviation_from_additivity = proportion_SR1 - (proportion_SS + proportion_R1R1) / 2
  
  # Calculate selection coefficient
  selection_coefficient = (proportion_R1R1 - proportion_SS) / proportion_SS
  
  # Calculate dominance coefficient
  dominance_coefficient = (proportion_SR1 - proportion_SS) / (proportion_R1R1 - proportion_SS)
  
  return(c(additivity = deviation_from_additivity, selection = selection_coefficient, dominance = dominance_coefficient))
}

# Set seed for reproducibility
set.seed(123)

bootstraps = 1000 

# Perform bootstrap with 1000 replicates
results.R1 = boot(data = data.viability.R1, statistic = bootstrap_function.R1, R = bootstraps)

# bootstrap values
deviation_additivity.R1 = results.R1$t0[1]
selection.R1 = results.R1$t0[2]
dominance.R1 = results.R1$t0[3]

# Extract bootstrap replicates (bootstrapped statistics)
boot_estimates = results.R1$t  # This extracts all the bootstrap replicates

# Calculate standard errors for each statistic
se_additivity_R1 = sd(boot_estimates[, 1])  # For additivity
se_selection_R1 = sd(boot_estimates[, 2])   # For selection coefficient
se_dominance_R1 = sd(boot_estimates[, 3])   # For dominance coefficient

# Compute confidence intervals
deviation_additivity_ci.R1 = boot.ci(boot.out = results.R1, type = "perc", index = 1)  # Deviation from additivity
selection_ci.R1 = boot.ci(boot.out = results.R1, type = "perc", index = 2)  # Selection coefficient
dominance_ci.R1 = boot.ci(boot.out = results.R1, type = "perc", index = 3)  # Dominance coefficient

# extract CI values
deviation_additivity_ci.R1.lower_ci = round(deviation_additivity_ci.R1$percent[4],2)  # Lower limit
deviation_additivity_ci.R1.upper_ci = round(deviation_additivity_ci.R1$percent[5],2)  # Upper limit

selection_ci.R1.lower_ci = round(selection_ci.R1$percent[4],2)  # Lower limit
selection_ci.R1.upper_ci = round(selection_ci.R1$percent[5],2)  # Upper limit

dominance_ci.R1.lower_ci = round(dominance_ci.R1$percent[4],2)  # Lower limit
dominance_ci.R1.upper_ci = round(dominance_ci.R1$percent[5],2)  # Upper limit


# Compute p-values for each statistic
pvalue_additivity.R1 = round(boot.pval(results.R1, index = 1, type = "perc", theta_null = 0),4)  # p-value for deviation from additivity
pvalue_selection.R1 = round(boot.pval(results.R1, index = 2, type = "perc", theta_null = 0),4)   # p-value for selection coefficient
pvalue_dominance.R1 = round(boot.pval(results.R1, index = 3, type = "perc", theta_null = 0.5),4)   # p-value for dominance coefficient
```

### 2.2 Ace R2 genotype

```{r}
# Data R2 
data.viability.R2 = data.viability %>% 
  # Filter data for Genotypes SS, SR2, and R2R2
  filter(Genotype %in% c("SS", "SR2", "R2R2")) %>%
  # Level data so the SS is the reference
  mutate(Genotype = factor(Genotype, levels = c("SS","SR2", "R2R2")))
  
# Fit logistic binomial model
model.R2 = glm(cbind(Emerge, Not_Emerge) ~ Genotype, data = data.viability.R2, family = "binomial")
  
# Calculate estimated marginal means for the Genotype on the response scale (probability) - otherwise the model gives logit values
emmeans(model.R2, ~ Genotype, type = "response")

# Extract values for table
emm.R2 = tidy(emmeans(model.R2, ~ Genotype, type = "response"))

SS.mean = round(emm.R2 %>% filter(Genotype == "SS") %>% pull(prob),2)
SR2.mean = round(emm.R2 %>% filter(Genotype == "SR2") %>% pull(prob),2)
R2R2.mean = round(emm.R2 %>% filter(Genotype == "R2R2") %>% pull(prob),2)

SS.se = round(emm.R2 %>% filter(Genotype == "SS") %>% pull(std.error),2)
SR2.se = round(emm.R2 %>% filter(Genotype == "SR2") %>% pull(std.error),2)
R2R2.se = round(emm.R2 %>% filter(Genotype == "R2R2") %>% pull(std.error),2)

# Perform pairwise comparisons
pairwise_comparisons.R2 = pairs(emmeans(model.R2, ~ Genotype, type = "response"))
  
# Adjusting for multiple comparisons using FDR (Benjamini-Hochberg)
pwc.R2 = summary(pairwise_comparisons.R2, adjust = "BH")
  
SS.SR2 = round(pwc.R2 %>% filter(contrast == "SS / SR2") %>% pull(odds.ratio),2)
SS.R2R2 = round(pwc.R2 %>% filter(contrast == "SS / R2R2") %>% pull(odds.ratio),2)
SR2.R2R2 = round(pwc.R2 %>% filter(contrast == "SR2 / R2R2") %>% pull(odds.ratio),2)

SS.SR2.se = round(pwc.R2 %>% filter(contrast == "SS / SR2") %>% pull(SE),2)
SS.R2R2.se = round(pwc.R2 %>% filter(contrast == "SS / R2R2") %>% pull(SE),2)
SR2.R2R2.se = round(pwc.R2 %>% filter(contrast == "SR2 / R2R2") %>% pull(SE),2)

SS.SR2.p = round(pwc.R2 %>% filter(contrast == "SS / SR2") %>% pull(p.value),4)
SS.R2R2.p = round(pwc.R2 %>% filter(contrast == "SS / R2R2") %>% pull(p.value),4)
SR2.R2R2.p = round(pwc.R2 %>% filter(contrast == "SR2 / R2R2") %>% pull(p.value),4)
```

```{r}
# Define the bootstrap function for R2
bootstrap_function.R2 = function(data, indices) {
  sample_data = data[indices, ]  # Resample data
  model = glm(cbind(Emerge, Not_Emerge) ~ Genotype, data = sample_data, family = "binomial")
  coefficients = coef(model)
  
  b0 = coefficients["(Intercept)"]
  b1 = coefficients["GenotypeSR2"]
  b2 = coefficients["GenotypeR2R2"]
  
  # Transform logit values to probabilities
  proportion_SS = exp(b0) / (1 + exp(b0))
  proportion_SR2 = exp(b0 + b1) / (1 + exp(b0 + b1))
  proportion_R2R2 = exp(b0 + b2) / (1 + exp(b0 + b2))
  
  # Calculate additivity
  deviation_from_additivity = proportion_SR2 - (proportion_SS + proportion_R2R2) / 2
  
  # Calculate selection coefficient
  selection_coefficient = (proportion_R2R2 - proportion_SS) / proportion_SS
  
  # Calculate dominance coefficient
  dominance_coefficient = (proportion_SR2 - proportion_SS) / (proportion_R2R2 - proportion_SS)
  
  return(c(additivity = deviation_from_additivity, selection = selection_coefficient, dominance = dominance_coefficient))
}

# Set seed for reproducibility
set.seed(123)

bootstraps = 1000 

# Perform bootstrap with 1000 replicates
results.R2 = boot(data = data.viability.R2, statistic = bootstrap_function.R2, R = bootstraps)

# Bootstrap values
deviation_additivity.R2 = results.R2$t0[1]
selection.R2 = results.R2$t0[2]
dominance.R2 = results.R2$t0[3]

# Extract bootstrap replicates (bootstrapped statistics)
boot_estimates = results.R2$t  # This extracts all the bootstrap replicates

# Calculate standard errors for each statistic
se_additivity_R2 = sd(boot_estimates[, 1])  # For additivity
se_selection_R2 = sd(boot_estimates[, 2])   # For selection coefficient
se_dominance_R2 = sd(boot_estimates[, 3])   # For dominance coefficient

# Compute confidence intervals
deviation_additivity_ci.R2 = boot.ci(boot.out = results.R2, type = "perc", index = 1)  # Deviation from additivity
selection_ci.R2 = boot.ci(boot.out = results.R2, type = "perc", index = 2)  # Selection coefficient
dominance_ci.R2 = boot.ci(boot.out = results.R2, type = "perc", index = 3)  # Dominance coefficient

# Extract CI values
deviation_additivity_ci.R2.lower_ci = round(deviation_additivity_ci.R2$percent[4],2)  # Lower limit
deviation_additivity_ci.R2.upper_ci = round(deviation_additivity_ci.R2$percent[5],2)  # Upper limit

selection_ci.R2.lower_ci = round(selection_ci.R2$percent[4],2)  # Lower limit
selection_ci.R2.upper_ci = round(selection_ci.R2$percent[5],2)  # Upper limit

dominance_ci.R2.lower_ci = round(dominance_ci.R2$percent[4],2)  # Lower limit
dominance_ci.R2.upper_ci = round(dominance_ci.R2$percent[5],2)  # Upper limit

# Compute p-values for each statistic
pvalue_additivity.R2 = round(boot.pval(results.R2, index = 1, type = "perc", theta_null = 0),4)  # p-value for deviation from additivity
pvalue_selection.R2 = round(boot.pval(results.R2, index = 2, type = "perc", theta_null = 0),4)   # p-value for selection coefficient
pvalue_dominance.R2 = round(boot.pval(results.R2, index = 3, type = "perc", theta_null = 0.5),4)   # p-value for dominance coefficient
```

### 2.3 Ace R3 genotype

```{r}
# Data R3 
data.viability.R3 = data.viability %>% 
  # Filter data for Genotypes SS, SR3, and R3R3
  filter(Genotype %in% c("SS", "SR3", "R3R3")) %>%
  # Level data so the SS is the reference
  mutate(Genotype = factor(Genotype, levels = c("SS","SR3", "R3R3")))
  
# Fit logistic binomial model
model.R3 = glm(cbind(Emerge, Not_Emerge) ~ Genotype, data = data.viability.R3, family = "binomial")
  
# Calculate estimated marginal means for the Genotype on the response scale (probability) - otherwise the model gives logit values
emmeans(model.R3, ~ Genotype, type = "response")

# Extract values for table
emm.R3 = tidy(emmeans(model.R3, ~ Genotype, type = "response"))

SS.mean = round(emm.R3 %>% filter(Genotype == "SS") %>% pull(prob),2)
SR3.mean = round(emm.R3 %>% filter(Genotype == "SR3") %>% pull(prob),2)
R3R3.mean = round(emm.R3 %>% filter(Genotype == "R3R3") %>% pull(prob),2)

SS.se = round(emm.R3 %>% filter(Genotype == "SS") %>% pull(std.error),2)
SR3.se = round(emm.R3 %>% filter(Genotype == "SR3") %>% pull(std.error),2)
R3R3.se = round(emm.R3 %>% filter(Genotype == "R3R3") %>% pull(std.error),2)

# Perform pairwise comparisons
pairwise_comparisons.R3 = pairs(emmeans(model.R3, ~ Genotype, type = "response"))
  
# Adjusting for multiple comparisons using FDR (Benjamini-Hochberg)
pwc.R3 = summary(pairwise_comparisons.R3, adjust = "BH")
  
SS.SR3 = round(pwc.R3 %>% filter(contrast == "SS / SR3") %>% pull(odds.ratio),2)
SS.R3R3 = round(pwc.R3 %>% filter(contrast == "SS / R3R3") %>% pull(odds.ratio),2)
SR3.R3R3 = round(pwc.R3 %>% filter(contrast == "SR3 / R3R3") %>% pull(odds.ratio),2)

SS.SR3.se = round(pwc.R3 %>% filter(contrast == "SS / SR3") %>% pull(SE),2)
SS.R3R3.se = round(pwc.R3 %>% filter(contrast == "SS / R3R3") %>% pull(SE),2)
SR3.R3R3.se = round(pwc.R3 %>% filter(contrast == "SR3 / R3R3") %>% pull(SE),2)

SS.SR3.p = round(pwc.R3 %>% filter(contrast == "SS / SR3") %>% pull(p.value),4)
SS.R3R3.p = round(pwc.R3 %>% filter(contrast == "SS / R3R3") %>% pull(p.value),4)
SR3.R3R3.p = round(pwc.R3 %>% filter(contrast == "SR3 / R3R3") %>% pull(p.value),4)
```

```{r}
# Define the bootstrap function for R3
bootstrap_function.R3 = function(data, indices) {
  sample_data = data[indices, ]  # Resample data
  model = glm(cbind(Emerge, Not_Emerge) ~ Genotype, data = sample_data, family = "binomial")
  coefficients = coef(model)
  
  b0 = coefficients["(Intercept)"]
  b1 = coefficients["GenotypeSR3"]
  b2 = coefficients["GenotypeR3R3"]
  
  # Transform logit values to probabilities
  proportion_SS = exp(b0) / (1 + exp(b0))
  proportion_SR3 = exp(b0 + b1) / (1 + exp(b0 + b1))
  proportion_R3R3 = exp(b0 + b2) / (1 + exp(b0 + b2))
  
  # Calculate additivity
  deviation_from_additivity = proportion_SR3 - (proportion_SS + proportion_R3R3) / 2
  
  # Calculate selection coefficient
  selection_coefficient = (proportion_R3R3 - proportion_SS) / proportion_SS
  
  # Calculate dominance coefficient
  dominance_coefficient = (proportion_SR3 - proportion_SS) / (proportion_R3R3 - proportion_SS)
  
  return(c(additivity = deviation_from_additivity, selection = selection_coefficient, dominance = dominance_coefficient))
}

# Set seed for reproducibility
set.seed(123)

bootstraps = 1000 

# Perform bootstrap with 1000 replicates
results.R3 = boot(data = data.viability.R3, statistic = bootstrap_function.R3, R = bootstraps)

# Bootstrap values
deviation_additivity.R3 = results.R3$t0[1]
selection.R3 = results.R3$t0[2]
dominance.R3 = results.R3$t0[3]

# Extract bootstrap replicates (bootstrapped statistics)
boot_estimates = results.R3$t  # This extracts all the bootstrap replicates

# Calculate standard errors for each statistic
se_additivity_R3 = sd(boot_estimates[, 1])  # For additivity
se_selection_R3 = sd(boot_estimates[, 2])   # For selection coefficient
se_dominance_R3 = sd(boot_estimates[, 3])   # For dominance coefficient

# Compute confidence intervals
deviation_additivity_ci.R3 = boot.ci(boot.out = results.R3, type = "perc", index = 1)  # Deviation from additivity
selection_ci.R3 = boot.ci(boot.out = results.R3, type = "perc", index = 2)  # Selection coefficient
dominance_ci.R3 = boot.ci(boot.out = results.R3, type = "perc", index = 3)  # Dominance coefficient

# Extract CI values
deviation_additivity_ci.R3.lower_ci = round(deviation_additivity_ci.R3$percent[4],2)  # Lower limit
deviation_additivity_ci.R3.upper_ci = round(deviation_additivity_ci.R3$percent[5],2)  # Upper limit

selection_ci.R3.lower_ci = round(selection_ci.R3$percent[4],2)  # Lower limit
selection_ci.R3.upper_ci = round(selection_ci.R3$percent[5],2)  # Upper limit

dominance_ci.R3.lower_ci = round(dominance_ci.R3$percent[4],2)  # Lower limit
dominance_ci.R3.upper_ci = round(dominance_ci.R3$percent[5],2)  # Upper limit

# Compute p-values for each statistic
pvalue_additivity.R3 = round(boot.pval(results.R3, index = 1, type = "perc", theta_null = 0),4)  # p-value for deviation from additivity
pvalue_selection.R3 = round(boot.pval(results.R3, index = 2, type = "perc", theta_null = 0),4)   # p-value for selection coefficient
pvalue_dominance.R3 = round(boot.pval(results.R3, index = 3, type = "perc", theta_null = 0.5),4)   # p-value for dominance coefficient
```

# 3. Table with statistics

```{r}
library(kableExtra)
library(htmltools)
source("./StatTable.R") # table for statistics
```

### 3.1 Ace R1 genotype

```{r}
# Call StatTable for model.R1
StatTable(model = model.R1, 
          directory = "./tables/", 
          display_title = "Effect of genotype (SS, SR1, R1R1) on viability",  # Title for the table
          save_title = "Viability.R1",  # Title for the saved file
           pred_labels = c("Intercept","SR1", "R1R1"), # replace with your labels
          dv_labels = "Genotype")

# dataframe with estimates
table.R1 = data.frame(
  Parameter = c("SS mean", "SR1 mean", "R1R1 mean", 
                "SS vs. SR1", "SS vs. R1R1", "SR1 vs. R1R1", 
                "SR1 - (SS + R1R1)/2 (additivity)", 
                "(R1R1 - SS) / R1R1 (selection, s)", "(SR1 - SS) / (R1R1 - SS) (dominance, h)"), 
  Viability = c(SS.mean, SR1.mean, R1R1.mean, 
                SS.SR1, SS.R1R1, SR1.R1R1, 
                deviation_additivity.R1, selection.R1, dominance.R1), 
  Std.Error = c(SS.se, SR1.se, R1R1.se, 
                SS.SR1.se, SS.R1R1.se, SR1.R1R1.se, 
                se_additivity_R1, se_selection_R1, se_dominance_R1),
  Lower.CI = c("-", "-", "-", "-", "-", "-",
                deviation_additivity_ci.R1.lower_ci,
                selection_ci.R1.lower_ci,
                dominance_ci.R1.lower_ci),
  Upper.CI = c("-", "-", "-", "-", "-", "-",
                 deviation_additivity_ci.R1.upper_ci,
                 selection_ci.R1.upper_ci,
                 dominance_ci.R1.upper_ci),
  P.value = c("-", "-", "-", SS.SR1.p, "<.0001", SR1.R1R1.p, 
         pvalue_additivity.R1, pvalue_selection.R1, pvalue_dominance.R1)
)

# Format the columns
table.R1$Viability = formatC(table.R1$Viability, format = "f", digits = 2)
table.R1$Std.Error = formatC(table.R1$Std.Error, format = "f", digits = 2)
table.R1$Lower.CI = formatC(table.R1$Lower.CI, format = "f", digits = 2)
table.R1$Upper.CI = formatC(table.R1$Upper.CI, format = "f", digits = 2)
table.R1$P.value = formatC(table.R1$P.value, format = "e", digits = 4)  # 4 decimals for p-values


table.R1.kable = kable(x = table.R1, digits = 4, align = 'l',
                             caption = "<b>Effect estimates from logistic model analysis for viability</b>") %>%
                       kable_classic(full_width = TRUE, html_font = "Times New Roman", font_size = 16) %>%
                       kable_styling(latex_options = c("striped", "hold_position"), position = "center") %>%
                       row_spec(0, bold = TRUE, font_size = 16) %>%  # Make header bold and set font size
                       row_spec(1:nrow(table.R1), font_size = 16)  # Smaller font for values

table.R1.kable
```

### 3.2 Ace R2 genotype

```{r}
# Call StatTable for model.R2
StatTable(model = model.R2, 
          directory = "./tables/", 
          display_title = "Effect of genotype (SS, SR2, R2R2) on viability",  # Title for the table
          save_title = "Viability.R2",  # Title for the saved file
          pred_labels = c("Intercept","SR2", "R2R2"), # replace with your labels
          dv_labels = "Genotype")

# dataframe with estimates
table.R2 = data.frame(
  Parameter = c("SS mean", "SR2 mean", "R2R2 mean", 
                "SS vs. SR2", "SS vs. R2R2", "SR2 vs. R2R2", 
                "SR2 - (SS + R2R2)/2 (additivity)", 
                "(R2R2 - SS) / R2R2 (selection, s)", "(SR2 - SS) / (R2R2 - SS) (dominance, h)"), 
  Viability = c(SS.mean, SR2.mean, R2R2.mean, 
                SS.SR2, SS.R2R2, SR2.R2R2, 
                deviation_additivity.R2, selection.R2, dominance.R2), 
  Std.Error = c(SS.se, SR2.se, R2R2.se, 
                SS.SR2.se, SS.R2R2.se, SR2.R2R2.se, 
                se_additivity_R2, se_selection_R2, se_dominance_R2),
  Lower.CI = c("-", "-", "-", "-", "-", "-",
               deviation_additivity_ci.R2.lower_ci,
               selection_ci.R2.lower_ci,
               dominance_ci.R2.lower_ci),
  Upper.CI = c("-", "-", "-", "-", "-", "-",
               deviation_additivity_ci.R2.upper_ci,
               selection_ci.R2.upper_ci,
               dominance_ci.R2.upper_ci),
  P.value = c("-", "-", "-", SS.SR2.p, "<.0001", "<.0001", 
              pvalue_additivity.R2, pvalue_selection.R2, pvalue_dominance.R2)
)

# Format the columns
table.R2$Viability = formatC(table.R2$Viability, format = "f", digits = 2)
table.R2$Std.Error = formatC(table.R2$Std.Error, format = "f", digits = 2)
table.R2$Lower.CI = formatC(table.R2$Lower.CI, format = "f", digits = 2)
table.R2$Upper.CI = formatC(table.R2$Upper.CI, format = "f", digits = 2)
table.R2$P.value = formatC(table.R2$P.value, format = "e", digits = 4)  # 4 decimals for p-values

table.R2.kable = kable(x = table.R2, digits = 4, align = 'l',
                             caption = "<b>Effect estimates from logistic model analysis for viability</b>") %>%
                       kable_classic(full_width = TRUE, html_font = "Times New Roman", font_size = 16) %>%
                       kable_styling(latex_options = c("striped", "hold_position"), position = "center") %>%
                       row_spec(0, bold = TRUE, font_size = 16) %>%  # Make header bold and set font size
                       row_spec(1:nrow(table.R2), font_size = 16)  # Smaller font for values

table.R2.kable
```

### 3.3 Ace R3 genotype

```{r}
# Call StatTable for model.R3
StatTable(model = model.R3, 
          directory = "./tables/", 
          display_title = "Effect of genotype (SS, SR3, R3R3) on viability",  # Title for the table
          save_title = "Viability.R3",  # Title for the saved file
          pred_labels = c("Intercept","SR3", "R3R3"), # replace with your labels
          dv_labels = "Genotype")

# dataframe with estimates
table.R3 = data.frame(
  Parameter = c("SS mean", "SR3 mean", "R3R3 mean", 
                "SS vs. SR3", "SS vs. R3R3", "SR3 vs. R3R3", 
                "SR3 - (SS + R3R3)/2 (additivity)", 
                "(R3R3 - SS) / R3R3 (selection, s)", "(SR3 - SS) / (R3R3 - SS) (dominance, h)"), 
  Viability = c(SS.mean, SR3.mean, R3R3.mean, 
                SS.SR3, SS.R3R3, SR3.R3R3, 
                deviation_additivity.R3, selection.R3, dominance.R3), 
  Std.Error = c(SS.se, SR3.se, R3R3.se, 
                SS.SR3.se, SS.R3R3.se, SR3.R3R3.se, 
                se_additivity_R3, se_selection_R3, se_dominance_R3),
  Lower.CI = c("-", "-", "-", "-", "-", "-",
               deviation_additivity_ci.R3.lower_ci,
               selection_ci.R3.lower_ci,
               dominance_ci.R3.lower_ci),
  Upper.CI = c("-", "-", "-", "-", "-", "-",
               deviation_additivity_ci.R3.upper_ci,
               selection_ci.R3.upper_ci,
               dominance_ci.R3.upper_ci),
  P.value = c("-", "-", "-", SS.SR3.p, "<.0001", SR3.R3R3.p, 
              pvalue_additivity.R3, pvalue_selection.R3, pvalue_dominance.R3)
)

# Format the columns
table.R3$Viability = formatC(table.R3$Viability, format = "f", digits = 2)
table.R3$Std.Error = formatC(table.R3$Std.Error, format = "f", digits = 2)
table.R3$Lower.CI = formatC(table.R3$Lower.CI, format = "f", digits = 2)
table.R3$Upper.CI = formatC(table.R3$Upper.CI, format = "f", digits = 2)
table.R3$P.value = formatC(table.R3$P.value, format = "e", digits = 4)  # 4 decimals for p-values

table.R3.kable = kable(x = table.R3, digits = 4, align = 'l',
                             caption = "<b> Effect estimates from logistic model analysis for viability</b>") %>%
                       kable_classic(full_width = TRUE, html_font = "Times New Roman", font_size = 16) %>%
                       kable_styling(latex_options = c("striped", "hold_position"), position = "center") %>%
                       row_spec(0, bold = TRUE, font_size = 16) %>%  # Make header bold and set font size
                       row_spec(1:nrow(table.R3), font_size = 16)  # Smaller font for values

table.R3.kable
```

## 3.4 R1+R2+R3 Table

```{r}
# remove first row in table 2 and table 3 because it is the same with table 1
table.R1.R2.R3 = rbind(table.R1, table.R2 %>% slice(-1), table.R3 %>% slice(-1))

# Use kable to create the table
table.R1.R2.R3.kable = table.R1.R2.R3 %>%
  kbl(caption = "Effect estimates from logistic model analysis for viability") %>%
  kable_classic(full_width = TRUE, html_font = "Arial") 
   
table.R1.R2.R3.kable

# Save the table as an HTML file
save_kable(table.R1.R2.R3.kable, file = "./tables/table_viability.html")

# Use webshot to convert HTML to PDF
webshot("./tables/table_viability.html", file = "./tables/table_viability.pdf")
```

## 3.5 Data for simulations

```{r}
table.viability = rbind (table.R1, table.R2, table.R3) %>%
  
  filter(Parameter == c("SS mean", "SR1 mean", "R1R1 mean", "SR2 mean", "R2R2 mean", "SR3 mean", "R3R3 mean")) %>%
  
  dplyr::select(Parameter, Viability)

write_csv(table.viability, "./data/model.viability.0ppm.csv")
```

# 4. Plots

### 4.1 Ace R1 genotype

```{r}
genotype_labels = c("SS","SR1","R1R1")

# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#30a4bc", "#ffce1b")
genotype_colors = c("#165094", "#7396BF","#ce8e58")

# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)



R1_viability = ggplot() +
  
  annotate("segment", x = 1, xend = 3, y = 89.4, yend = 76.4, colour = "#ce8e58", size=0.5, alpha=0.5) +
  
 
  # This plots individual values for each genotype
  geom_jitter(data = data.viability %>% filter(Genotype %in% c("SS", "SR1", "R1R1")),
             aes(x=Genotype, y = Survival, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +
  
   # add a line for SS mean
  geom_hline(yintercept = 89.4, color = "black", linetype = "solid", size = 0.25) +
  
  # This plots the mean viability for each genotype
  geom_point(data = data.viability.summary %>% filter(Genotype %in% c("SS","SR1", "R1R1")),
              aes(x=Genotype, y = mean, color = Genotype, shape = Genotype), size = 8, alpha = 1) +
  
  # This plots the mean and standard error for each genotype
  geom_errorbar(data = data.viability.summary %>% filter(Genotype %in% c("SS", "SR1", "R1R1")),
                aes(x=Genotype, ymin = mean - se, ymax = mean + se, color = Genotype), width = 0) +
  

  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
  # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R1
  geom_text(aes(x = 1.2, y = 120, 
                label = "s = -0.15 +/-0.03*** \nh = 0.29 +/-0.2"), 
            size = 8, hjust = 0, vjust = 1) +
  
  
  scale_y_continuous(breaks = seq(40, 120, 10), 
                     limits = c(40, 120)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Viability (%)") +
  
  
  theme_classic() +
  
  theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")
    
  

R1_viability

ggsave("./plots/R1_viability.png",R1_viability, width = 5 , height = 5)
```

### 4.2 Ace R2 genotype

```{r}
genotype_labels = c("SS","SR2","R2R2")

# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#30a4bc", "#ec4176")
genotype_colors = c("#165094", "#165094","#e1615e")


# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)


R2_viability = ggplot() +
  
  annotate("segment", x = 1, xend = 3, y = 89.4, yend = 74.0, colour = "#e1615e", size=0.5, alpha=0.5) +
  
 
  # This plots individual values for each genotype
  geom_jitter(data = data.viability %>% filter(Genotype %in% c("SS", "SR2", "R2R2")),
             aes(x=Genotype, y = Survival, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +
  
   # add a line for SS mean
  geom_hline(yintercept = 89.4, color = "black", linetype = "solid", size = 0.25) +
  
  # This plots the mean viability for each genotype
  geom_point(data = data.viability.summary %>% filter(Genotype %in% c("SS","SR2", "R2R2")),
              aes(x=Genotype, y = mean, color = Genotype, shape = Genotype), size = 8, alpha = 1) +
  
  # This plots the mean and standard error for each genotype
  geom_errorbar(data = data.viability.summary %>% filter(Genotype %in% c("SS", "SR2", "R2R2")),
                aes(x=Genotype, ymin = mean - se, ymax = mean + se, color = Genotype), width = 0) +
  

  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
  # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R2
  geom_text(aes(x = 1.2, y = 120, 
                label = "s = -0.17 +/-0.05*** \nh = 0.17 +/-0.15*"), 
            size = 8, hjust = 0, vjust = 1) +
  
  
  scale_y_continuous(breaks = seq(40, 120, 10), 
                     limits = c(40, 120)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Viability (%)") +
  
  
  theme_classic() +
  
  theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")
    
  

R2_viability

ggsave("./plots/R2_viability.png",R2_viability, width = 5, height = 5)
```

### 4.3 Ace R3 genotype

```{r}

genotype_labels = c("SS","SR3","R3R3")

# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#97D1DD",  "#782e6e")
genotype_colors = c("#165094", "#165094", "#702771")

# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)


R3_viability = ggplot() +
  
  annotate("segment", x = 1, xend = 3, y = 89.4, yend = 77.6, colour = "#782e6e", size=0.5, alpha=0.5) +
  
 
  # This plots individual values for each genotype
  geom_jitter(data = data.viability %>% filter(Genotype %in% c("SS", "SR3", "R3R3")),
             aes(x=Genotype, y = Survival, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +
  
   # add a line for SS mean
  geom_hline(yintercept = 89.4, color = "black", linetype = "solid", size = 0.25) +
  
  # This plots the mean viability for each genotype
  geom_point(data = data.viability.summary %>% filter(Genotype %in% c("SS","SR3", "R3R3")),
              aes(x=Genotype, y = mean, color = Genotype, shape = Genotype), size = 8, alpha = 1) +
  
  # This plots the mean and standard error for each genotype
  geom_errorbar(data = data.viability.summary %>% filter(Genotype %in% c("SS", "SR3", "R3R3")),
                aes(x=Genotype, ymin = mean - se, ymax = mean + se, color = Genotype), width = 0) +
  

  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
  # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R2
  geom_text(aes(x = 1.2, y = 120, 
                label = "s = -0.13 +/-0.02*** \nh = 0.32 +/-0.19"), 
            size = 8, hjust = 0, vjust = 1) +
  
  
  scale_y_continuous(breaks = seq(40, 120, 10), 
                     limits = c(40, 120)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Viability (%)") +
  
  
  theme_classic() +
  
  theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")
    
  

R3_viability

ggsave("./plots/R3_viability.png",R3_viability, width = 5 , height = 5)
```

# 4. References

1.  Tutorial for 1-Way ANOVAL:

    <https://www.datanovia.com/en/lessons/anova-in-r/#post-hoc-tests>

    <https://whitlockschluter3e.zoology.ubc.ca/Tutorials%20using%20R/R_tutorial_Normal_and_sample_means.html>

2.  Regression with categorical variable:

    <https://www.youtube.com/watch?v=9yTui_LoSOc>

    <http://www.sthda.com/english/articles/40-regression-analysis/163-regression-with-categorical-variables-dummy-coding-essentials-in-r/>
