---
title: "Comparison of fecundity between Ace genotypes in the absence of malathion"
author: "Marianna Karageorgi"
date: "2025-05-07"
---

# 1. Data

Experiment was conducted from July 17- 20, 2023.

We introduced five 3-day-old females of each genotype in egg-laying bottles fitted with egg-laying caps filled with Drosophila growth media. Each day over a period of 3 days we counted the number of eggs laid. Each of the genotypes were tested in 8-10 replicate egg-laying bottles. The egg-laying caps were examined under a stereoscope where we counted the eggs laid each day over the 3-day period.

The bottles were randomized and the experimenter, Anastasia Andreeva, was blind to the genotype in each bottle.

### Data wrangling

```{r}
library(tidyverse)
library(data.table)
```

```{r}
data = fread("./data/Fecundity_FINAL.csv")

fecundity.data = data %>%
  mutate(eggs = rowSums(data[, 5:7], na.rm = TRUE), 
         Total = 50 ,
         eggs_laid = eggs / Total * 100 ) %>%
  filter(Nr_flies_Experiment_End == 5) %>%
  dplyr::select(Genotype, eggs_laid) %>%
  slice(-c(60, 63)) %>% # experimental error - wrong nr of flies in egg lay cages
  mutate(Genotype = factor(Genotype, levels = c("SS", "SR1","R1R1", "SR2", "R2R2","SR3", "R3R3"))) # Convert Genotype to a factor
```

### Summary Stats

```{r}
data.fecundity.summary = group_by(fecundity.data, Genotype) %>%
  summarise(
    count = n(),
    mean = mean(eggs_laid, na.rm = TRUE),
    sd = sd(eggs_laid, na.rm = TRUE),
    se = sd / sqrt(8)
  )
```

### Fecundity data: Exploratory Data Analysis

```{r}

# Define the custom colors for the palette
genotype_colors = c("black", "#FCCC1A", "#FCCC1A","#8601AF", "#8601AF","#FE2712",    "#FE2712")

# Define the linetypes for the boxes
linetypes = c(1, 2, 2, 2, 1, 1, 1)

# Use the chosen colors and linetypes in your plot
fecundity_plot = ggplot(fecundity.data, aes(Genotype, eggs_laid)) +
  geom_boxplot(aes(linetype = Genotype, color = Genotype), fill = NA) +
  geom_point(aes(color = Genotype)) +  # Add a color aesthetic here
  scale_color_manual(values = genotype_colors) +  # Map the colors to the genotypes
  scale_linetype_manual(values = linetypes) +
  ylim(0, 350)+
  labs(x = NULL, y = "Eggs Laid") +
  theme_classic() +
  theme(legend.position = "right") + 
  
  theme(legend.position = "none")

fecundity_plot
```

# 2. Model data

Use linear regression model to analyze fecundity data for homozygous sensitive SS, homozygous resistant RR, and heterozygous SR genotypes for each resistant allele.

The model is:

eggs=b0​+b1​⋅SR1+b2​⋅R1R1+b3​⋅SR2+b4​⋅R2R2+ b5​⋅SR3+b6​⋅R3R3 ϵ

The expected number of eggs for each genotype is represented:

-   For SS (reference group): eggsSS = b0

-   For SR1: eggsSR1 = b0 + b1

-   For R1R1: eggsR1R1 = b0 + b2

-   For SR2: eggsSR2 = b0 + b3

-   For R2R2: eggsR2R2 = b0 + b4

-   For SR3: eggsSR3 = b0 + b5

-   For R3R3: eggsR3R3 = b0 + b6

```{r}
# Fit linear model
model.R = lm(eggs_laid ~ Genotype, data = fecundity.data)

summary(model.R)
```

# 3. Metrics & Bootstrap for fecundity

Estimate difference and metric parameters using coefficients from the linear model. Then use bootstrapping to estimate confidence intervals for each parameter.

**Additivity: Comparison between SR vs. (SS+RR)/2**

Additivity is calculated as the difference between the SR genotype and the average of SS and RR genotypes. It can be compared to the expected value under additivity.

-   Expected value under additivity: SR- (SS+RR)/2

-   Significance testing : H0 = the estimate for additivity is 0. Ha1 = the estimate for dominance is higher than 0 Ha2 = the estimate for dominance is lower than 0.

**Selection coefficient: Estimate ratio (RR - SS)/ SS**

The selection coefficient is the ratio of the difference between RR and SS genotypes to the SS genotype. This parameter indicates how selection affects the fitness of the resistant genotype compared to the sensitive genotype.

-   Estimate ratio (RR-SS)/SS

-   Significance testing with two-sided t-test: H0 = the estimate for selection equals to 0. Ha = the estimate for selection is different from 0.

**Dominance coefficient: Estimate (SR−SS)/(RR−SS) (h)**

The dominance coefficient measures the degree to which the heterozygous genotype (SR) exhibits fitness that is intermediate between the homozygous genotypes (SS and RR).

-   Estimate ratio (SR-SS)/(RR-SS)

-   Significance testing with two one-sided t-test: H0 = the estimate for dominance is 0.5. Ha1 = the estimate for dominance is higher than 0.5. Ha2 = the estimate for dominance is lower than 0.5

```{r}
library(boot)
library(broom)

# bootstrap & estimate metrics

# bootstrap function
metrics.boot = function(data, original) { 
  # Create bootstrap sample
  d = data %>%
    group_by(Genotype) %>%
    # randomly select eight rows from each Genotype group, sampling with replacement
    slice_sample(n = 8, replace = TRUE) %>% # n = number of reps in each fecundity group
    ungroup()
  
  # Fit model to bootstrap sample 
  model = lm(eggs_laid ~ Genotype, data = d)
  
  # Extract coefficients to calculate metrics for each boot
  coefficients = model$coefficients  # Fix: extract coefficients from model
  b0 = coefficients["(Intercept)"]
  b1 = coefficients["GenotypeSR1"]
  b2 = coefficients["GenotypeR1R1"]
  b3 = coefficients["GenotypeSR2"]
  b4 = coefficients["GenotypeR2R2"]
  b5 = coefficients["GenotypeSR3"]
  b6 = coefficients["GenotypeR3R3"]
  
  # Get values for each genotype to calculate metrics for each boot
  ss   = b0
  sr1  = b0+b1
  r1r1 = b0+b2
  sr2  = b0+b3
  r2r2 = b0+b4
  sr3  = b0+b5
  r3r3 = b0+b6
  
  # Calculate all metrics for each R allele for each boot
  # For R1
  ss_sr1_diff = -b1
  ss_r1r1_diff = -b2
  sr1_r1r1_diff = b1 - b2
  dev.add.r1 = b1 - b2/2
  s.r1 = b2/b0
  h.r1 = b1/b2
  
  # For R2
  ss_sr2_diff = -b3
  ss_r2r2_diff = -b4
  sr2_r2r2_diff = b3 - b4
  dev.add.r2 = b3 - b4/2
  s.r2 = b4/b0
  h.r2 = b3/b4
  
  # For R3
  ss_sr3_diff = -b5
  ss_r3r3_diff = -b6
  sr3_r3r3_diff = b5 - b6
  dev.add.r3 = b5 - b6/2
  s.r3 = b6/b0
  h.r3 = b5/b6
  
  # Return with explicit names
  result_vector = c(r1r1, r2r2, r3r3, sr1, sr2, sr3, ss, 
                   dev.add.r1, dev.add.r2, dev.add.r3,
                   s.r1, s.r2, s.r3,
                   h.r1, h.r2, h.r3,
                   ss_sr1_diff, ss_sr2_diff, ss_sr3_diff,
                   ss_r1r1_diff, ss_r2r2_diff, ss_r3r3_diff,
                   sr1_r1r1_diff, sr2_r2r2_diff, sr3_r3r3_diff)
  
  names(result_vector) = c("R1R1", "R2R2", "R3R3","SR1", "SR2", "SR3", "SS",
                          "DevAdd_R1", "DevAdd_R2", "DevAdd_R3",
                          "S_R1", "S_R2", "S_R3",
                          "H_R1", "H_R2", "H_R3",
                          "diff_SS.SR1", "diff_SS.SR2", "diff_SS.SR3",
                          "diff_SS.R1R1", "diff_SS.R2R2", "diff_SS.R3R3",
                          "diff_SR1.R1R1", "diff_SR2.R2R2", "diff_SR3.R3R3")
  return(result_vector)
}

# run the bootstrap
set.seed(123)
nboot=1000
result = boot(data = fecundity.data,
               statistic = metrics.boot,
               R = nboot)  # R for number of bootstraps

result$t0 # original estimate based on data
result$t  # boot samples estimates

summary(result)
tidy(result)
```

```{r}
### Fecundity ###

# Report the original estimate (result$t0) based on the data  
# Use the 2.5% and 97.5% quantiles of the bootstrap distribution (result$t) for the 95% ci

# Process bootstrap results into a single LD50 table
fecundity.results = function(result) {
  # Define genotypes and indices
  genotypes = c("R1R1", "R2R2", "R3R3", "SR1", "SR2", "SR3", "SS")
  fecundity_indices = 1:7
  
  # Process fecundity values
  fecundity_table = do.call(rbind, lapply(1:length(genotypes), function(i) {
    # Extract values
    original = result$t0[i]
    boot_samples = result$t[, i]
    ci = quantile(boot_samples, c(0.025, 0.975))
    
    data.frame(
      Genotype = genotypes[i],
      Estimate = original,
      Lower.CI = ci[1],
      Upper.CI = ci[2]
    )
  }))
  
  return(fecundity_table)
}

# Get the viability data frame
fecundity = fecundity.results(result)
fecundity
```

```{r}
### DIFFERENCES & METRICS ####

# Report the original estimate (result$t0) based on the data  
# Use the 2.5% and 97.5% quantiles of the bootstrap distribution (result$t) for the 95% ci
# Report the p-value calculated from the bootstrap distribution

# First, define the p.value function
p.value = function(boot_samples, null_value) {
  p_less = mean(boot_samples <= null_value, na.rm = TRUE)
  p_greater = mean(boot_samples >= null_value, na.rm = TRUE)
  return(2 * min(p_less, p_greater)) # which tail is more extreme
}

# Then define metrics.results
metrics.results = function(result) {
  # Define alleles
  alleles = c("R1", "R2", "R3")
  
  # Define all metrics
  all_metrics = list(
    
    # Derived metrics
    list(type = "DevAdd", name = "Additivity", indices = 8:10, null_value = 0),
    list(type = "S", name = "Selection coefficient", indices = 11:13, null_value = 0),
    list(type = "H", name = "Dominance coefficient", indices = 14:16, null_value = 0.5),
    
    # Derived differences
    list(type = "diff_SS.SR", name = "SS-SR difference", indices = 17:19, null_value = 0),
    list(type = "diff_SS.RR", name = "SS-RR difference", indices = 20:22, null_value = 0),
    list(type = "diff_SR.RR", name = "SR-RR difference", indices = 23:25, null_value = 0)
  )
  
  # Process metrics
  metrics_table = do.call(rbind, lapply(1:length(all_metrics), function(m) {
    metric = all_metrics[[m]]
    
    do.call(rbind, lapply(1:length(alleles), function(a) {
      idx = metric$indices[a]
      original = result$t0[idx]
      boot_samples = result$t[, idx]
      ci = quantile(boot_samples, c(0.025, 0.975), na.rm = TRUE)
      p_val = p.value(boot_samples, metric$null_value)
      
      data.frame(
        Metric = metric$name,
        Allele = alleles[a],
        Estimate = original,
        Lower.CI = ci[1],
        Upper.CI = ci[2],
        p.value = p_val
      )
    }))
  }))
  
  # Return the formatted table
  return(metrics_table)
}

# Now call metrics.results
metrics = metrics.results(result)
metrics[metrics$p.value==0,]$p.value= 1/nboot # this is so that p.value is not 0
```

# 3. Table for fecundity

```{r}
# combine dataframes for the final table
fecundity.format = fecundity %>%
  rename(Genotype.Allele = Genotype) %>%
  mutate(Metric = "fecundity") %>%  
  dplyr::select(Genotype.Allele, Metric, Estimate, Lower.CI, Upper.CI) %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  select(Metric, Genotype.Allele, Estimate, Lower.CI, Upper.CI)

metrics.format =  metrics %>%
  rownames_to_column(var = "column") %>% separate(column, into = c("Metric", "Genotype.Allele"), sep = "_") %>%
  dplyr::select(-Allele) %>%
  mutate(across(where(is.numeric), ~round(., 3)))

all.metrics = bind_rows(fecundity.format, metrics.format) %>%
  mutate(Genotype.Allele = factor(Genotype.Allele, 
                                  levels = c("SS", "SR1", "SR2", "SR3", "R1R1", "R2R2", "R3R3", 
                                             "SS.SR1","SS.SR2","SS.SR3", "SS.R1R1","SS.R2R2","SS.R3R3", "SR1.R1R1","SR2.R2R2","SR3.R3R3",
                                             "R1", "R2", "R3")),
         Metric = factor(Metric, 
                         levels = c("fecundity","diff","DevAdd", "S", "H"))) %>%
  arrange(Metric, Genotype.Allele)
```

```{r}
library(kableExtra)
library(htmltools)
library(webshot)
```

```{r}
all.metrics.table = all.metrics %>%
  unite(col = "Parameter", Genotype.Allele, Metric, sep = "_") %>%
  rename(Fecundity = Estimate) %>%
  
  mutate(Parameter = case_when(
    
    Parameter == "SS_fecundity" ~ "SS",
    Parameter == "SR1_fecundity" ~ "SR1",
    Parameter == "R1R1_fecundity" ~ "R1R1",
    Parameter == "SR2_fecundity" ~ "SR2",
    Parameter == "R2R2_fecundity" ~ "R2R2",
    Parameter == "SR3_fecundity" ~ "SR3",
    Parameter == "R3R3_fecundity" ~ "R3R3",

    Parameter == "SS.SR1_diff" ~ "SS vs. SR1",
    Parameter == "SS.R1R1_diff" ~ "SS vs. R1R1",
    Parameter == "SR1.R1R1_diff" ~ "SR1 vs. R1R1",
    Parameter == "SS.SR2_diff" ~ "SS vs. SR2",
    Parameter == "SS.R2R2_diff" ~ "SS vs. R2R2",
    Parameter == "SR2.R2R2_diff" ~ "SR2 vs. R2R2",
    Parameter == "SS.SR3_diff" ~ "SS vs. SR3",
    Parameter == "SS.R3R3_diff" ~ "SS vs. R3R3",
    Parameter == "SR3.R3R3_diff" ~ "SR3 vs. R3R3",

    Parameter == "R1_DevAdd" ~ "SR1 - (SS + R1R1)/2 (additivity)",
    Parameter == "R2_DevAdd" ~ "SR2 - (SS + R2R2)/2 (additivity)",
    Parameter == "R3_DevAdd" ~ "SR3 - (SS + R3R3)/2 (additivity)",

    Parameter == "R1_S" ~ "(R1R1 - SS) / SS (selection, s)",
    Parameter == "R2_S" ~ "(R2R2 - SS) / SS (selection, s)",
    Parameter == "R3_S" ~ "(R3R3 - SS) / SS (selection, s)",

    Parameter == "R1_H" ~ "(SR1 - SS) / (R1R1 - SS) (dominance, h)",
    Parameter == "R2_H" ~ "(SR2 - SS) / (R2R2 - SS) (dominance, h)",
    Parameter == "R3_H" ~ "(SR3 - SS) / (R3R3 - SS) (dominance, h)")) %>%
   #replace NA with -
  mutate(across(everything(), ~replace(., is.na(.), "-")))


table.kable = kable(x = all.metrics.table, digits = 3, align = 'l',
                       caption = "<b>Effect estimates from linear model analysis for fecundity</b>",
                    row.names = FALSE) %>%
                       kable_classic(full_width = TRUE, html_font = "Times New Roman", font_size = 16) %>%
                       kable_styling(latex_options = c("striped", "hold_position"), position = "center") %>%
                       row_spec(0, bold = TRUE, font_size = 16) %>%  # Make header bold and set font size
                       row_spec(1:nrow(all.metrics.table), font_size = 16)  # Smaller font for values

    


table.kable

# Save the table as an HTML file
save_kable(table.kable, file = "./tables/table_fecundity.html")

# Use webshot to convert HTML to PDF
webshot("./tables/table_fecundity.html", file = "./tables/table_fecundity.pdf")
```

## Percent changes between genotypes

Calculate changes between genotypes for each allele

e.g. for R1:

(SR1-SS)/SS x 100

(R1R1-SS)/ SS x100

```{r}
all.metrics.calc = all.metrics %>%
  unite(col = "Parameter", Genotype.Allele, Metric, sep = "_") %>%
  rename(fecundity = Estimate) %>%
  
  mutate(fecundity.SS = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "fecundity") %>% pull(Estimate)) %>%

  # Then calculate the change
  mutate(change = (fecundity - fecundity.SS)/fecundity.SS * 100)
```

# 4. Plots

### 4.1 Ace R1 genotype

```{r}
library(ggtext)

genotype_labels = c("SS","SR1","R1R1")

# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#97D1DD", "#ffce1b")
genotype_colors = c("#165094", "#7396BF","#ce8e58")


# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)


R1_fecundity = ggplot() +
  
  annotate("segment", x = 1, xend = 3, 
           y = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "fecundity") %>% pull(Estimate), 
           yend = all.metrics %>% filter(Genotype.Allele == "R1R1" & Metric == "fecundity") %>% pull(Estimate), 
           colour = "#ce8e58", size=0.5, alpha=0.5) +
  
  # add a line for SS mean
  geom_hline(yintercept = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "fecundity") %>% pull(Estimate), 
             color = "black", linetype = "solid", size = 0.25) +
  
  # This plots all individual values for each genotype
  geom_jitter(data = fecundity.data %>% filter(Genotype %in% c("SS", "SR1", "R1R1")),
             aes(x=Genotype, y = eggs_laid, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +

  
  # This plots fecundity for each genotype
  geom_point(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR1", "R1R1")),
              aes(x=Genotype.Allele, y = Estimate, color = Genotype.Allele, shape = Genotype.Allele), size = 8, alpha = 1) +

  # This plots the confidence interval for each genotype
  geom_errorbar(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR1", "R1R1")),
                aes(x=Genotype.Allele, ymin = Lower.CI, ymax = Upper.CI, 
                    color = Genotype.Allele), width = 0) +
  
  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
   # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R1
  geom_richtext(aes(x = 0.7, y = 320, 
                label = "*s* = -0.28 (-0.47, -0.15)** <br> *h* = 0.71 (-0.05, 1.32)"), 
            size = 8, hjust = 0, vjust = 1, 
            fill = NA, label.color = NA) +

  scale_y_continuous(breaks = seq(50, 320, 50), 
                     limits = c(50,320)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Fecundity") +
  
  theme_classic() +
  
   theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")

R1_fecundity

ggsave("./plots/R1_fecundity.png",R1_fecundity,  width = 5 , height = 5)
```

### 4.2 Ace R2 genotype

```{r}

genotype_labels = c("SS","SR2","R2R2")



# Define the custom colors for each genotype
#genotype_colors = c("#30a4bc", "#97D1DD", "#ec4176")
genotype_colors = c("#165094", "#7396BF","#e1615e")


# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)


R2_fecundity = ggplot() +
  
  annotate("segment", x = 1, xend = 3, 
           y = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "fecundity") %>% pull(Estimate), 
           yend = all.metrics %>% filter(Genotype.Allele == "R2R2" & Metric == "fecundity") %>% pull(Estimate), 
           colour = "#ec4176", size=0.5, alpha=0.5) +
  
  # add a line for SS mean
  geom_hline(yintercept = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "fecundity") %>% pull(Estimate), 
             color = "black", linetype = "solid", size = 0.25) +
  
  # This plots all individual values for each genotype
  geom_jitter(data = fecundity.data %>% filter(Genotype %in% c("SS", "SR2", "R2R2")),
             aes(x=Genotype, y = eggs_laid, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +

  
  # This plots the mean fecundity for each genotype
  geom_point(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR2", "R2R2")),
              aes(x=Genotype.Allele, y = Estimate, color = Genotype.Allele, shape = Genotype.Allele), size = 8, alpha = 1) +

  # This plots the confidence interval for each genotype
  geom_errorbar(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR2", "R2R2")),
                aes(x=Genotype.Allele, ymin = Lower.CI, ymax = Upper.CI, 
                    color = Genotype.Allele), width = 0) +
  
  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
   # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R2
  geom_richtext(aes(x = 0.7, y = 320, 
                label = "*s* = -0.35 (-0.32, 0.03) <br> *h* = 0.40 (-2.48, 3.09)"), 
            size = 8, hjust = 0, vjust = 1, 
            fill = NA, label.color = NA) +
  
  scale_y_continuous(breaks = seq(50, 320, 50), 
                     limits = c(50,320)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Fecundity") +
  
  theme_classic() +
  
   theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")

R2_fecundity

ggsave("./plots/R2_fecundity.png",R2_fecundity ,width = 5 , height = 5)
```

### 4.3 Ace R3 genotype

```{r}

genotype_labels = c("SS","SR3","R3R3")

# Define the custom colors for each genotype
#genotype_colors = c("black", "#4d4d4d",  "#FE2712")
#genotype_colors = c("#30a4bc", "#30a4bc",  "#782e6e")
genotype_colors = c("#165094", "#165094", "#702771")



# Define the shapes for each genotype
genotype_shapes = c(16, 15, 17)


R3_fecundity = ggplot() +
  
  annotate("segment", x = 1, xend = 3, 
           y = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "fecundity") %>% pull(Estimate), 
           yend = all.metrics %>% filter(Genotype.Allele == "R3R3" & Metric == "fecundity") %>% pull(Estimate), 
           colour = "#782e6e", size=0.5, alpha=0.5) +
  
  # add a line for SS mean
  geom_hline(yintercept = all.metrics %>% filter(Genotype.Allele == "SS" & Metric == "fecundity") %>% pull(Estimate), 
             color = "black", linetype = "solid", size = 0.25) +
  
  # This plots all individual values for each genotype
  geom_jitter(data = fecundity.data %>% filter(Genotype %in% c("SS", "SR3", "R3R3")),
             aes(x=Genotype, y = eggs_laid, color = Genotype, shape = Genotype), size = 4, alpha = 0.4, width = 0.1) +

  
  # This plots the mean fecundity for each genotype
  geom_point(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR3", "R3R3")),
              aes(x=Genotype.Allele, y = Estimate, color = Genotype.Allele, shape = Genotype.Allele), size = 8, alpha = 1) +

  # This plots the confidence interval for each genotype
  geom_errorbar(data = all.metrics %>% filter(Genotype.Allele %in% c("SS","SR3", "R3R3")),
                aes(x=Genotype.Allele, ymin = Lower.CI, ymax = Upper.CI, 
                    color = Genotype.Allele), width = 0) +
  
  # Map the colors to the genotypes
  scale_color_manual(values = genotype_colors) +
  
   # Map the shapes to the genotypes
  scale_shape_manual(values = genotype_shapes) +
  
  # Add the s and h coefficients for R3
  geom_richtext(aes(x = 0.7, y = 320, 
                label = "*s* = -0.43 (-0.60, -0.31)*** <br> *h* = 0.08 (-0.45, 0.27)***"), 
            size = 8, hjust = 0, vjust = 1, 
            fill = NA, label.color = NA) +
  

  scale_y_continuous(breaks = seq(50, 320, 50), 
                     limits = c(50,320)) +
  
  scale_x_discrete(labels = genotype_labels) +
  
  labs(x = "", y = "Fecundity") +
  
  theme_classic() +
  
   theme(
     panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
 #   panel.background= element_rect(color ="#FCCC1A", size = 2),
    
    # Increase the size of the axis titles
    axis.title.y = element_text(size = 24),

    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 20),
 
    # No legend
    legend.position = "none")

R3_fecundity

ggsave("./plots/R3_fecundity.png",R3_fecundity,  width = 5 , height = 5)
```

# 5. References

1.  Regression with categorical variable:

    <https://www.youtube.com/watch?v=9yTui_LoSOc>

    <http://www.sthda.com/english/articles/40-regression-analysis/163-regression-with-categorical-variables-dummy-coding-essentials-in-r/>

2.  Add p values manually:

    <http://rpkgs.datanovia.com/ggpubr/reference/stat_pvalue_manual.html#examples>

3.  Bootstrap:

    <https://stats.oarc.ucla.edu/r/faq/how-can-i-generate-bootstrap-statistics-in-r/>

    <https://advstats.psychstat.org/book/ci/bootstrap.php>
