---
title: "Comparison of malathion LD50 between Ace genotypes"
author: "Marianna Karageorgi"
date: "2023-06-24"
output: html_document
---

# 1. Data

Starting July 4th 2023, eggs of each D.melanogaster genotype (SS, R1R1, R2R2, R3R3, SR1, SR2, SR3) were exposed to malathion at a range of concentrations diluted in regular Drosophila media (2 trials, 50 eggs/trial).

• dose_ppm = ppm concentration of malathion

• Total = number of eggs exposed per trial

• Alive = number of eggs that survived to adulthood

• Dead = number of eggs that did not survive to adulthood

### Data Wrangling

```{r}
library(tidyverse)
library(data.table)
```

```{r}
viability.range_malathion = fread("./data/Resistance_Data_FINAL.csv")

resistance.data = viability.range_malathion %>%
  
  #create columns alive, total, and dead
  mutate(Alive = rowSums(viability.range_malathion[, 6:15], na.rm = TRUE), 
         Total = 50, 
         Dead = Total - Alive) %>%
    
  mutate(Survival = Alive / Total * 100) %>%
    
  dplyr::rename(Dose = `Dose_(ppm)`) %>%
  
  dplyr::select(Genotype, Replicate, Dose, Alive, Dead, Total, Survival) 
  
resistance.data$Genotype = factor(resistance.data$Genotype,
                                     levels=c("SS","SR1","SR2","SR3","R1R1", "R2R2", "R3R3"))
```

### Summary stats

```{r}
resistance.data.summary = resistance.data %>%
  
  # group replicates 
  group_by(Genotype, Dose) %>%
  
  #summary statistics
  dplyr::summarize(mean_Survival = mean(Survival), 
                   replicates = n(), sd = sd(Survival), 
                   # se = sd / sqrt(2), # i'm not calculating se because size of groups is unequal 
                   .groups = "keep") 
```

### Dose-Response Data: Exploratory Data Analysis

```{r}
library(ggpubr) 
library(ggthemes)

# Define the custom colors for each genotype
genotype_labels = c("S/S", "S/R1", "R1/R1", "S/R2","R2/R2", "S/R3", "R3/R3")
  
# Define the custom colors for the palette
genotype_colors = c("black", "#FCCC1A","#8601AF","#FE2712",  "#FCCC1A", "#8601AF",   "#FE2712")

# Define the linetypes for the λινεσ
linetypes = c(1, 2, 2, 2, 1, 1, 1)

# Define the shapes for each genotype
genotype_shapes = c(16, 1, 16, 1, 16, 1, 16)

resistance_plot = ggplot(data = resistance.data.summary, 
                        aes(x = Dose, y = mean_Survival, color = Genotype, linetype = Genotype)) + 
  
  #this plots for mean of Survival at each dose per Genotype
  geom_line(aes(x = Dose, y = mean_Survival), size = 1, alpha = 1) +
   
  #this plots for standard error of Survival at each dose per Genotype
  geom_errorbar(aes(ymin=mean_Survival-sd, ymax=mean_Survival + sd), width = 0.2) +
  
  # Map the colors to the genotypes
  scale_color_manual(name = "Ace genotype",
                    labels = genotype_labels,
                    values = genotype_colors) +
  
  # Map the linetypes to the genotypes
  scale_linetype_manual(name = "Ace genotype",
                        labels = genotype_labels,
                        values = linetypes) +

  
  scale_x_continuous(breaks=seq(0, 30, 2.5),
                     limits = c(0, 30)) +
  
  scale_y_continuous(breaks=seq(0, 100, 20)) +
  
  labs(x= "Dose (ppm)", y = "Egg to adult viability (%)") +
  
  theme_classic() +
  theme(
    # Increase the size of the axis titles
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    
    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 13, angle = 90, vjust = 0.5),
    axis.text.y = element_text(size = 15),
    
     # Position the legend inside the plot in the upper left hand corner
    legend.position = c(0.99, 0.99),
        legend.background = element_rect(fill = "transparent"),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(3, 3, 3, 3),
        legend.text.align = 0,
        legend.text = element_text(size = 10)
  ) 
  

resistance_plot
```

# 2. Dose response analysis

```{r}
library(drc)
library(broom)

library(kableExtra)
library(htmltools)
library(webshot)
```

-   Fit logistic dose-response curves to the resistance data of each Ace genotype. The drm() function is the key function in the drc package for fitting dose-response curves. Survival is the response variable (y-axis) and Dose is the dose (x-axis). Genotype is the classification variable. data= resistance.data identifies the name of the data file. fct=L.4() is the argument for the logistic curve with four parameters. Here, we fixed two of the parameters, the lower and upper limits, at 0 and 100, respectively. Notice that the drm() function does not produce an output. All the information on the model fit is stored in the object under our defined name resistance.data.model. This will be read by further commands.

-   The summary() function provides a summary of the parameter estimates for the resistance.data.model. In this case, we have constrained the lower and upper limits to 0 and 100, respectively for all datasets. The parameters that are estimated are the slope (b) and the effective dose (e) and their standard errors. The p-values tell us if the parameters are different from zero.

-   The ED() function calculates the estimated effective dose for the resistance data based on the fitted model

### Model data for all genotypes

```{r}
# data for all R genotypes
resistance.data

# Model data for R genotype
model.R = drm(Survival ~ Dose, Genotype,
              data = resistance.data, 
              fct = L.4(fixed = c(NA, 0, 100, NA)))

# Summary of the model
ld50.summary = tidy(model.R)

# Extract ED50 values using the ED() function
ed50_values = ED(model.R, c(50))
```

### Table for LD50 for each genotype

```{r}
# dataframe with estimates
table.Ld50 = ld50.summary %>%
  filter(term == "e") %>%
  rename(Genotype = curve, LD50 = estimate) %>%
  dplyr::select(-term)

# Format the columns
table.Ld50$LD50 = formatC(table.Ld50$LD50, format = "f", digits = 2)
table.Ld50$std.error = formatC(table.Ld50$std.error, format = "f", digits = 2)
table.Ld50$statistic = formatC(table.Ld50$statistic, format = "f", digits = 2)  
table.Ld50$p.value = formatC(table.Ld50$p.value, format = "e", digits = 4)  # 4 decimals for scientific notation


# Use kable to create the table
table.Ld50.kable = table.Ld50 %>%
  kbl(caption = "Dose response curve analysis for each Ace genotype") %>%
  kable_classic(full_width = TRUE, html_font = "Arial") 
   

table.Ld50.kable

# Save the table as an HTML file
save_kable(table.Ld50.kable, file = "./tables/table_drc_LD50.html")

# Use webshot to convert HTML to PDF
webshot("./tables/table_drc_LD50.html", file = "./tables/table_drc_LD50.pdf")
```

### **All Pairwise curve comparisons**

Use a global F test to assess if the parameters of the curves for the two different Ace genotypes are identical. Based on these two models we can calculate an approximate F-test. Then use fdr correction.

```{r}
# Fit a model that does not distinguish among two curves (Model 1). Fit a model which distinguish the dose response curve for each of them (Model 2). Based on these two models then calculate an approximate F-test. Do it for all pairwise comparisons:SS-SR1,SS-R1R1, SR1-R1R1

# Define the pairs of genotypes to compare
genotype_pairs = list(
  c("SS", "SR1"), c("SS", "R1R1"), c("SS", "SR2"), c("SS", "R2R2"), c("SS", "SR3"), c("SS", "R3R3"), c("SR1", "R1R1"), c("SR1", "SR2"), c("SR1", "R2R2"), c("SR1", "SR3"), c("SR1", "R3R3"), c("R1R1", "SR2"), c("R1R1", "R2R2"), c("R1R1", "SR3"), c("R1R1", "R3R3"), c("SR2", "R2R2"), c("SR2", "SR3"), c("SR2", "R3R3"), c("R2R2", "SR3"), c("R2R2", "R3R3"), c("SR3", "R3R3")
)

# Initialize an empty list to store the results
anova_results = list()

# Loop through each pair of genotypes
for (pair in genotype_pairs) {
  # Subset the data for the current pair of genotypes
  subset_data = subset(resistance.data, Genotype %in% pair)
  
  # Fit Model 1: does not distinguish between genotypes
  model1 = drm(Survival ~ Dose, 
                data = subset_data,
                fct = L.4(fixed = c(NA, 0, 100, NA)))
  
  # Fit Model 2: distinguishes between genotypes
  model2 = drm(Survival ~ Dose, Genotype,
                data = subset_data, 
                fct = L.4(fixed = c(NA, 0, 100, NA)))
  
  # Calculate F test
  test_result = tidy(anova(model1, model2))
  
  # Add column to identify the genotype pair being compared
  test_result = test_result %>%
    mutate(Comparison = paste(pair[1], "vs", pair[2]))
  
  # Append the result to the list
  anova_results = c(anova_results, list(test_result))
  
}

# Combine all pairwise comparisons into a single data frame
anova_results.R = do.call(rbind, anova_results)

# perform fdr correction
f.test.fdr = anova_results.R %>%
  
  # Filter for the "2nd model" term
  filter(term == "2nd model") %>%
  
  # Apply FDR correction to the p.value column
  mutate(fdr_corrected = p.adjust(p.value, method = "fdr"))

f.test.fdr
```

### **Predict egg-adult viability using model**

```{r}
# grid of doses and genotypes for which we want predictions
new_doses = expand.grid(Dose = seq(0, 30, by = 0.2), Genotype = unique(resistance.data$Genotype))

# predict survival using the resistance data model
predictions.R = predict(model.R, newdata = new_doses, type = "response")

# attach the predictions to the new_doses data frame
new_doses$Survival = predictions.R
```

### **Predict egg-adult viability at 2.5ppm and 7.5ppm for simulations**

```{r}
# Specify the doses for which predictions are needed
doses = c(2.5, 5, 7.5, 10, 15)

# Create a grid of these specific doses and genotypes for which we want predictions
specific.doses = expand.grid(Dose = doses, Genotype = unique(resistance.data$Genotype))

# Predict survival using the resistance data model
predictions.specific.doses.R = predict(model.R, newdata = specific.doses, type = "response")

# Attach the predictions to the specific_doses data frame
specific.doses$Survival = round(predictions.specific.doses.R,2)

specific.doses 

write_csv(specific.doses, "./data/drc.model.viability.malathion.csv")
```

### **Plot dose response curves**

```{r}
genotype_labels = c(
  expression("S/S"),
  expression("S/R1"),
  expression("S/R2"),
  expression("S/R3"),
  expression("R1/R1"),
  expression("R2/R2"),
  expression("R3/R3")
  )

# Define the custom colors for each genotype
genotype_colors = c("#165094", "#E2BB9B" ,"#EDA09E","#A97DAA","#ce8e58", "#e1615e",  "#702771")


# Define the linetypes for the lines
linetypes = c(1, 2, 2, 2, 1, 1, 1)

# Define the shapes for each genotype
genotype_shapes = c(16, 15, 15,15, 17,17, 17)

Ace.resistance.curves.R = ggplot(data = resistance.data,
       aes(x = Dose, y = Survival, group= Genotype, color= Genotype, linetype = Genotype)) + 
  
  # Annotate to indicate viability at 2.5 and 7.5ppm
annotate("rect", xmin=2.3, xmax=2.7, ymin=-Inf, ymax=Inf, alpha=0.8, fill="#e5e5e5") +
annotate("rect", xmin=7.3, xmax=7.7, ymin=-Inf, ymax=Inf, alpha=0.8, fill="#e5e5e5") +

  geom_point(aes(color=Genotype, shape = Genotype), size = 2) + 
  
  #plot model predictions
  geom_line(data= new_doses, aes(x = Dose, y = Survival, color = Genotype, linetype = Genotype), size = 1) +
  
  
  # Map the colors to the genotypes
  scale_color_manual(name = "Ace genotype",
                    labels = genotype_labels,
                    values = genotype_colors) +
  
  # Map the linetypes to the genotypes
  scale_linetype_manual(name = "Ace genotype",
                        labels = genotype_labels,
                        values = linetypes) +
  
  # Map the shapes to the genotypes
  scale_shape_manual(name = "Ace genotype",
                        labels = genotype_labels,
                        values = genotype_shapes) +

  scale_x_continuous(breaks=seq(0, 30, 2.5),
                     limits = c(-0.5, 30)) +
  
  scale_y_continuous(breaks=seq(0, 100, 10),
                     limits = c(-5,100)) +
  
  labs(title= "", x = "Dose (ppm)",  y = "Egg to adult viability (%)") + 
  
  theme_classic() +
  
  theme(
    # Increase the size of the axis titles
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    
    # Increase the size of the axis tick labels
    axis.text.x = element_text(size = 13, angle = 90, vjust = 0.5),
    axis.text.y = element_text(size = 15),
    
     # Position the legend inside the plot in the upper left hand corner
    legend.position = c(0.99, 0.99),
        legend.background = element_rect(fill = "transparent"),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(3, 3, 3, 3),
        legend.text.align = 0,
        legend.text = element_text(size = 10)
  ) 


Ace.resistance.curves.R
ggsave("./plots/Ace.resistance.curves.R.png", Ace.resistance.curves.R, width = 6, height = 4)
```

# 3. References

1.  Drc package and the logistic model:

    <https://doseresponse.github.io/drc/reference/logistic.html>
